---
title: "Intro to MCMC: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_22_Assignment_Key.Rmd") -->

#### Due by the beginning of next class (11/23)

Save this Rmarkdown script as a file named “YourlastName_Assignment22.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment22.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about Bayesian inference

2. extrapolate from in-class exercises to apply Bayesian inference to arbitrary distributions

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

1. Explain, in your own words, the goal and steps of an MCMC algorithm.

**The goal of an MCMC is to characterize the posterior probability distribution 
of a model given some data by sampling from each position in parameter space 
in proportion to its posterior probability**

The steps of an MCMC are as follows:

1. Initiate the chain with a randomly chosen parameter value
2. Calculate the posterior probability of the model given the data 
(the log likelihood of the data given that parameter value plus the prior probability of that parameter value)
3. Propose an update to the current parameter value
4. Calculate the posterior probability of the updated model given the data
5. Calculate the ratio R of the proposed and current posterior probabilities
6. Accept the proposed move with a probability proportional to R

Repeat steps 3-6 many times.

### Question 2

Below is a chunk of R code that implements an MCMC algorithm 
to estimate the rate parameter ($\lambda$) of a Poisson distribution. 
This code is uncommented - please write a comment for each line 
explaining what that line is doing.

```{r}

# function for calculating the posterior probability of the model given the data
calcPostProb <- function(poisData,lambda){
	# calcualte poisson log likelihood
	lnL <- sum(dpois(x=poisData,lambda=lambda,log=TRUE))
	# calculate log prior probability
		# prior is normal with mean 1 and large variance
	lnPrior <- dnorm(x=lambda,mean=1,sd=1e6,log=TRUE)
	# sum log likelihood and log prior to get log posterior probability, and return
	return(lnL + lnPrior)
}

# function for running an MCMC to infer the rate parameter of a Poisson distribution
mcmcPoisson <- function(poisData, nSteps){
	# initiate empty vector of lambda samples
	lambdaSamples <- rep(NA,nSteps)
	# initiate empty vector of posterior probability
	postProbSamples <- rep(NA,nSteps)
	# initiate chain with randomly chosen value for lambda
	lambda <- abs(rnorm(1))
	# calculate posterior probability of model given the data
	postProb <- calcPostProb(poisData,lambda)
	# log the initial value of lambda
	lambdaSamples[1] <- lambda
	# log the initial value of the posterior probability
	postProbSamples[1] <- postProb
	# do a for loop over the remaining nSteps-1 iterations of the MCMC
	for(i in 1:nSteps){
		# propose an update to lambda
		lambdaPrime <- lambda + rnorm(n=1,mean=0,sd=0.3)
		# lambda is constrained to be >0, so if the proposed value of lambda is negative, reject the move
		if(lambdaPrime > 0){
			# calculate the posterior probability with the proposed value of lambda
			postProbPrime <- calcPostProb(poisData,lambdaPrime)
			# calculate the ratio of the posterior probabilities of the proposed and current model
			#	note that because we've calculated the log posterior probabilities, 
			#		their ratio is the exponent of the difference between them
			postRatio <- exp(postProbPrime - postProb)
			# accept the propose move with a probability proportional to the ratio of the posterior probabilities
			if(postRatio > runif(n=1,min=0,max=1)){
				# if the move is accepted, the proposed value of lambda becomes the current value
				lambda <- lambdaPrime
				# if the move is accepted, the proposed value of the posterior probability becomes the current value
				postProb <- postProbPrime
			}
		}
		# log the current value of lambda
		lambdaSamples[i] <- lambda
		# log the current value of the posterior probability
		postProbSamples[i] <- postProb
	}
	# return a list containing the estimates of lambda
	# and the corresponding posterior probabilities
	return(
		list(
			"lambdaSamples" = lambdaSamples,
			"postProbSamples" = postProbSamples)
	)
}
```


### Question 3

Use the code above to do inference on the rate parameter 
I used to simulate the Poisson-distributed data saved in the 
.Robj file in the Class 22 homework folder on D2L.

Provide some visualizations of the results of your MCMC run

```{r}
load("poisData.Robj")
run <- mcmcPoisson(poisData,1e3)
layout(matrix(c(1,2,5,5,3,4,5,5),nrow=2,ncol=4,byrow=TRUE))
	plot(run$postProbSamples,xlab="MCMC iterations",ylab="posterior probability")
	plot(run$lambdaSamples,xlab="MCMC iterations",ylab="estimates of lambda")
	plot(run$postProbSamples[201:1e3],xlab="MCMC iterations post burnin",ylab="posterior probability")
	plot(run$lambdaSamples[201:1e3],xlab="MCMC iterationspost burnin",ylab="estimates of lambda")

	hist(run$lambdaSamples[201:1e3],xlab="estimates of lambda",main="marginal posterior distribution of lambda\n(post burn-in)")
		credInt <- quantile(run$lambdaSamples[201:1e3],c(0.025,0.975))
		abline(v=credInt,col=2,lty=2,lwd=2)
		legend(x="topleft",legend=c("95% credible interval","true value"),lty=c(2,1),lwd=c(1,2),col=c(2,4))
		abline(v=5.5,col="blue",lwd=3)
```



