---
title: "Intro to MCMC: Breakout room"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_22__Assignment_breakoutroom_exercise.Rmd") -->

Here are the functions we defined in class for getting a prior, log-likelihood, and posterior. You'll want to run the code in this block to store the functions.

```{r}
# define a prior log probability function (we are coin novices)
ln.prior <- function(p){
	return(dunif(x=p, min=0, max=1, log=TRUE))
}

# define a log-likelihood function
ln.L <- function(flips,n,p){
	return(dbinom(x=flips, size=n, prob=p, log=TRUE))
}

# define a posterior probability function
ln.post <- function(flips,n,p){
	return(ln.L(flips,n,p) + ln.prior(p)) 
  ##note that we can add because it's logs
}
```


Below is a function to run an MCMC that takes the following inputs:
n.iter = the number of iterations that you want your MCMC to run
flips = the number of heads in n flips that you observe
n = the number of flips

The function returns all the samples from the posterior from the MCMC.

```{r}

binom.mcmc <- function(n.iter,flips,n){
	p.samples <- rep(NA,n.iter) #make an empty vector for the samples
	p <- runif(1) #set the prior to be uniform
	post <- ln.post(flips,n,p) #calculate the posterior
	for(i in 1:n.iter){
		p.prime <- p + rnorm(1,0,0.05)
		post.prime <- -Inf
		if(is.finite(ln.prior(p.prime))){
			post.prime <- ln.post(flips,n,p.prime)
		}
		if(exp(post.prime-post) > runif(1)){
			p <- p.prime
		}
		p.samples[i] <- p
	}
	return(p.samples)
}

```

You're given a new coin and you want to know if it's fair. You flip the coin 20 times and get heads 3 times. Use binom.mcmc to make an MCMC with 50 iterations to to estimate the posterior distribution of p(heads) for flipping a coin 20 times and getting heads 3 times. What is the mean value of p(heads) from your MCMC interations?
\
\

Try it again with 500 iterations. Then with 1000 iterations.\
\

How many iterations do you need to get a good estimate of the probability of heads?
