---
title: "Intro to MCMC: Assignment"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_22_Assignment.Rmd") -->

#### Due by midnight before the beginning of next class (11/17)

Save this Rmarkdown script as a file named “YourlastName_Assignment22.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment22.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about Bayesian inference

2. extrapolate from in-class exercises to apply Bayesian inference to arbitrary distributions

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

1. Explain, in your own words, the goal and steps of an MCMC algorithm.

The goal of an MCMC algorithm is to estimate a parameter describing a dataset. This is done by randomly sampling a parameter, testing the posterior probability of the data given that parameter, and then resampling a nearby paramater to see if this new parameter has a higher probability. Over many steps the algorithm can describe the posterior distribution and find the best parameter estimate.

### Question 2

Below is a chunk of R code that implements an MCMC algorithm 
to estimate the rate parameter ($\lambda$) of a Poisson distribution. 
This code is uncommented - please write a comment for each line 
explaining what that line is doing.

```{r}

poisData = rpois(500, 5)
lambda = 2

#function that takes the data and an estimate of Lambda
calcPostProb <- function(poisData,lambda){
  ##calculate the log likelihood of poisData given the estimate of Lambda
	lnL <- sum(dpois(x=poisData,lambda=lambda,log=TRUE))
	# calculate the likeligood of the prior data
	lnPrior <- dnorm(x=lambda,mean=1,sd=1e6,log=TRUE)
	#sum them together (multiply but it's logs) to get the numerator of the Bayes equation
	return(lnL + lnPrior) 
}

nSteps = 100

## a function to run th mcmc
mcmcPoisson <- function(poisData, nSteps){
  ## make a vector to stick samples in
	lambdaSamples <- rep(NA,nSteps)
	## make a vector to stick the posterior probabilities in 
	postProbSamples <- rep(NA,nSteps)
	#sample an initial lambda
	lambda <- abs(rnorm(1))
	#calculate the posterior probability for that lambda
	postProb <- calcPostProb(poisData,lambda)
	# stick the lambda in lambda samples, and stick the posterior prob there as well
	lambdaSamples[1] <- lambda
	postProbSamples[1] <- postProb
	## now for each step
	for(i in 1:nSteps){
	    ## get a new lambda sample
	  	lambdaPrime <- lambda + rnorm(n=1,mean=0,sd=0.3)
	  	## i th esample is greater than 0
		if(lambdaPrime > 0){
		  ## calculate the posterior prob for th esample
			postProbPrime <- calcPostProb(poisData,lambdaPrime)
			## calculate R 
			postRatio <- exp(postProbPrime - postProb)
			## if it's good
			if(postRatio > runif(n=1,min=0,max=1)){
			  ## update to the new lambdas and posterior prime samples
				lambda <- lambdaPrime
				postProb <- postProbPrime
			}
		}
	  	## save the values
		lambdaSamples[i] <- lambda
		postProbSamples[i] <- postProb
	}
	return(
		list(
			"lambdaSamples" = lambdaSamples,
			"postProbSamples" = postProbSamples)
	)
}
```


### Question 3

Use the code above to do inference on the rate parameter 
I used to simulate the Poisson-distributed data saved in the 
.Robj file in the Class 22 homework folder on D2L.

Provide some visualizations of the results of your MCMC run.
Use your imagination here - if you were a data scientist, 
what plots would you want to show your client to convince 
them you had done a good job?

```{r}

load('poisData.Robj')

myMCMC = mcmcPoisson(poisData, 500)

plot(myMCMC$lambdaSamples, xlab = "steps", ylab = "lambda estimate")
plot(myMCMC$postProbSamples, xlab = "steps", ylab = "posterior probability")

hist(myMCMC$lambdaSamples, xlab = "lambda", col = "darkgray", border="white", breaks=100)

plot(myMCMC$lambdaSamples, myMCMC$postProbSamples, xlab = "lambda", ylab = "posterior probability", bty="n")

```



      