---
title: "rethinking_hw1"
output:
html_document:default
pdf_document: default
date: "2022-08-15"
---


1. Suppose the globe tossing data (Chapter 2) had turned out to be 4 water and 11 land. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book.

```{r}

#define grid
pgrid <- seq(from=0, to=1, length.out=20)

#define prior
prior <- rep(1,20)

#compute likeligood for the grid
likelihood <- dbinom(4, size=15, prob=pgrid)

unstd.posterior <- likelihood*prior

posterior <- unstd.posterior/sum(unstd.posterior)

plot(pgrid,posterior)

```


Now suppose the data are 4 water and 2 land. Compute the posterior
again, but this time use a prior that is zero below p = 0.5 and a constant
above p = 0.5. This corresponds to prior information that a majority of the
Earth’s surface is water.

```{r}
newprior <- c(rep(0, length.out=10), rep(2,length=10))

likelihood <- dbinom(4, size=6, prob=pgrid)
unstd.posterior <- likelihood*newprior
posterior <- unstd.posterior/sum(unstd.posterior)
plot(pgrid,posterior)


```