---
title: "rethinking_hw1"
output:
html_document:default
pdf_document: default
date: "2022-08-15"
---


1. Suppose the globe tossing data (Chapter 2) had turned out to be 4 water and 11 land. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book.

```{r}

#define grid
pgrid <- seq(from=0, to=1, length.out=1000)

#define prior
prior <- rep(1,1000)

#compute likeligood for the grid
likelihood <- dbinom(4, size=15, prob=pgrid)

unstd.posterior <- likelihood*prior

posterior <- unstd.posterior/sum(unstd.posterior)

plot(pgrid,posterior)

```


Now suppose the data are 4 water and 2 land. Compute the posterior
again, but this time use a prior that is zero below p = 0.5 and a constant
above p = 0.5. This corresponds to prior information that a majority of the
Earth’s surface is water.

```{r}
newprior <- c(rep(0, length.out=500), rep(2,length=500))

likelihood <- dbinom(4, size=6, prob=pgrid)
unstd.posterior <- likelihood*newprior
posterior <- unstd.posterior/sum(unstd.posterior)
plot(pgrid,posterior)


```

For the posterior distribution from 2, compute 89% percentile and HPDI
intervals. Compare the widths of these intervals. Which is wider? Why? If
you had only the information in the interval, what might you misunderstand
about the shape of the posterior distribution?


```{r}
mysamples <- sample(pgrid, prob=posterior, size=1e4, replace=T)

hist(mysamples, xlim=c(0,1))

quantile(samples(c(0.)))

```

