---
title: "rethinking_hw2"
output:
html_document:default
pdf_document: default
date: "2022-08-22"
---


Construct a linear regression of weight as predicted by height, using the adults (age 18 or greater) from the Howell1 dataset. The heights listed below were recorded in the !Kung census, but weights were not recorded for these individuals. Provide predicted weights and 89% compatibility intervals for each of these individuals. That is, fill in the table below, using model-based predictions.

```{r}
library(rethinking)

data( Howell1 )

#rename to make code easier
d <- Howell1

#get just the adults
d2 <- d[ d$age >= 18 , ]

#get the means
Hbar <- mean(d2$height)

#make a list of the data
dat <- list(W=d2$weight,H=d2$height,Hbar=Hbar)

#run the model
m1 <- quap(
    alist(
        W ~ dnorm( mu , sigma ) ,
        mu <- a + b*( H - Hbar ) ,
        a ~ dnorm( 60 , 10 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 10 )
    ) , data=dat )

#make the table
dat2 <- list( H=c(140,160,175,100) , Hbar=Hbar )

# simulate the data using the  using the rethinking package
h_sim <- sim( m1 , data=dat2 )
Ew <- apply(h_sim,2,mean)
h_ci <- apply(h_sim,2,PI,prob=0.89)

datr <- cbind( H=c(140,160,175) , Ew , L89=h_ci[1,] , U89=h_ci[2,] )
round(datr,1)



```


From the Howell1 dataset, consider only the people younger than 13 year old. Estimate the causal association between age and weight. Assume that age influences weight through two paths. First, age influences height, and height influences weight. Second, age directly influences weight through age-related changes in muscle growth and body proportions. All of this implies this causal model (DAG):

Use a linear regression to estimate the total (not just direct) causal effect of each year of growth on weight. Be sure to carefully consider the priors. Try using prior predictive simulation to assess what they imply.

```{r}

## pull out under thirteens
d3 <- d[ d$age < 13 , ]

# sim from priors
n <- 10
a <- rnorm(n,5,1)
b <- rlnorm(n,0,1)
# blank(bty="n")
plot( NULL , xlim=range(d3$age) , ylim=range(d3$weight) )
for ( i in 1:n ) abline( a[i] , b[i] , lwd=3 , col=2, xlab = 'age', ylab = 'weight')


#make a list of the data
dat <- list(W=d3$weight,H=d3$height,A=d3$age,Abar=mean(d3$age))

#run the model
m2 <- quap(
    alist(
        W ~ dnorm( mu , sigma ) , #prior for weight
        mu <- a + b*A , # formula for predicting weight from age
        a ~ dnorm( 5 , 1 ) , # prior for intercept
        b ~ dlnorm( 0 , 1 ) , #prior for relationship
        sigma ~ dexp(1) #noise
    ) , data=dat)

precis(m2)


plot( d3$age , d3$weight , lwd=3, col=2 )
post <- extract.samples(m2)
for ( i in 1:10 ) abline( post$a[i] , post$b[i] , lwd=3 , col=1 )

```


Now suppose the causal association between age and weight might be different for boys and girls. Use a single linear regression, with a categorical variable for sex, to estimate the total causal effect of age on weight separately for boys and girls. How do girls and boys differ? Provide one or more posterior contrasts as a summary.


```{r}
dat <- list(W=d3$weight,H=d3$height,A=d3$age,S=d3$male+1)


m3 <- quap(
    alist(
        W ~ dnorm( mu , sigma ),
        mu <- a[S] + b[S]*A,
        a[S] ~ dnorm(5,1),
        b[S] ~ dlnorm(0,1),
        sigma ~ dexp(1)
    ), data=dat )

precis(m3, depth=2)

plot( d$age , d$weight , lwd=3, col=ifelse(d$male==1,4,2) , xlab="age (years)" , ylab="weight (kg)" )
Aseq <- 0:12

# girls
muF <- link(m3,data=list(A=Aseq,S=rep(1,13)))
shade( apply(muF,2,PI,0.99) , Aseq , col=col.alpha(2,0.5) )
lines( Aseq , apply(muF,2,mean) , lwd=3 , col=2 )

# boys
muM <- link(m3,data=list(A=Aseq,S=rep(2,13)))
shade( apply(muM,2,PI,0.99) , Aseq , col=col.alpha(4,0.5) )
lines( Aseq , apply(muM,2,mean) , lwd=3 , col=4 )


# contrast at each age
Aseq <- 0:12

#simulate the mean based on the model
mu1 <- sim(m3,data=list(A=Aseq,S=rep(1,13))) #female
mu2 <- sim(m3,data=list(A=Aseq,S=rep(2,13))) #male
mu_contrast <- mu1
for ( i in 1:13 ) mu_contrast[,i] <- mu2[,i] - mu1[,i]
plot( NULL , xlim=c(0,13) , ylim=c(-15,15) , xlab="age" , ylab="weight difference (boys-girls)" )

for ( p in c(0.5,0.67,0.89,0.99) )
shade( apply(mu_contrast,2,PI,prob=p) , Aseq )

abline(h=0,lty=2,lwd=2)

for ( i in 1:13 ) points( mu_contrast[1:1000,i] , col=ifelse(mu_contrast[1:1000,i]>0,4,2) , lwd=3 )

```