---
title: "Intro to Probability 2: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment.Rmd") -->

#### Due by the beginning of next class (11/2)

Save this Rmarkdown script as a file named “YourlastName_Assignment16.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment16.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Define the following terms: 

1a) random variable
 
 Some variable that is the outcome of a random/stochastic process
 \
 \
 \
 
1b) probability distribution function
 
 This function describes the probability of outcomes across the sample space
 \
 \
 \


### Question 2
\
<div class="centered">
$\Large {{n}\choose{k}}p^k (1-p)^{n-k}$
</div>

2a) Using your own words, break down the equation above and 
describe what each part means (or is doing).

This equation describes the probability distribution function of coin flipping (or similar processes).
n is the number of trials (or flips)
k is the number of successes
and p is the probability of success

It is the combination of two smaller expression.
$p^{k} (1-p)^{n-k}$ is the probability of getting k successes and n-k failures given a certain underlying probability k. This expression is the product of all the probabilities of success (p) and all the probabilities of failure (1-p).
${n}\choose{k}$ is the number of permutations that will give you k successes

2b) Using the function of this equation that you coded up in class, calculate the probability of 
seeing 71 successes out of 531 trials, where the probability of success is 0.12345.

```{r, echo=T}
coinFlipProb <- function(n,k,p){
  myprob = choose(n,k)*p^k*(1-p)^(n-k)
  return(myprob)
}

coinFlipProb(531,71,.12345)

```

### Question 3

October in the Northern Hemisphere marks the start of flu season.
If our class were meeting in person, what is the probability 
that no one in the classroom would have an active case of the flu?

\
\

For the purposes of this calculation, assume the following:

-there are 55 people in our class\
-each of us can be treated as a random draw from Ingham County\
-there are ~280,000 people in Ingham County\
-over the last 5 days, the average per-day number of newly detected flu cases in Ingham County is 40\
-someone infected with the flu is infectious for 5 days\
-only 1 in every 3 flu infections is detected.\

3a) What is the analytical probability that no one in our classroom 
has an active flu infection?

First -- how many people have the flu?
40 per day * 5 days * 3 cases per detected case = 600
P(Infection in Ingham County) = 600/280000 = 0.002142857
P(Not infected in Ingham County) = 1 - 600/280000 = 0.9978571

P(No one has flu) = P(Each person doesn't have the flu)^(Number of people)
= (0.9978571)^55
= 0.8887085

3b) What is the probability at least one person in the room **_does_** have an active 
flu infection

P(Someone has the flu) = 1 - P(no one has the flu) = 0.1112915

3c) What is probability 2 people in the room are infected?

Now we can use our equation from above

```{r, echo=T}

coinFlipProb(55,2,0.002142857)

```



3d) Write a function for simulating this problem for gatherings of arbitrary size.
The arguments of the function should be the number of people in the room 
and the probability that a randomly selected person in the county has an active flu infection,
and it should return the probability (from simulations), 
that no one in the room has an active flu infection.

```{r, echo=T}
myN = 55
myP = 0.002142857

fluClass <- function(myN, myP){
  myRuns = sapply(1:1e4,function(x){
  myclass = sample(c(1,0),myN, replace=T, prob = c(myP, 1-myP))
  mysick = sum(myclass) > 0
  return(mysick)})
  probFlu = sum(myRuns)/1e4
  return(probFlu)
}

fluClass(55, 0.002142857)

```


