---
title: "Intro to Probability 2: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment_Key.Rmd") -->

#### Due by the beginning of next class (11/2)

Save this Rmarkdown script as a file named “YourlastName_Assignment16.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment16.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Define the following terms: 

1a) random variable
 
 \
**the outcome of a stochastic process**
 \
 
1b) probability distribution function
 
 \
**a rule or formula that describes the probability of all possible outcomes of a stochastic process**
 \


### Question 2

<div class="centered">
$\Large {{n}\choose{k}}p^k (1-p)^{n-k}$
</div>

2a) Using your own words, break down the equation above and 
describe what each part means (or is doing).

\
**The first part ("n choose k") describes the number of ways to get _k_ successes 
out of _n_ trials.
The second part ($p^k \times (1-p)^{n-k})$) gives the probability of getting _k_ successes
out of _n_ trials, where the probability of a success is _p_.
Together, these two parts describe the probability of seeing _k_ successes out of _n_ trials, 
where the probability of a success is _p_.
\

2b) Using the function of this equation that you coded up in class, calculate the probability of 
seeing 71 successes out of 531 trials, where the probability of success is 0.12345.

\
`r choose(531,71) * 0.12345^71 * (1-0.12345)^(531-71)`
**0.03947493**
\


### Question 3

\
\

October in the Northern Hemisphere marks the start of flu season.
If our class were meeting in person, what is the probability 
that no one in the classroom would have an active case of the flu?
\

For the purposes of this calculation, assume the following:

-there are 55 people in our class\
-each of us can be treated as a random draw from Ingham County\
-there are ~280,000 people in Ingham County\
-over the last 5 days, the average per-day number of newly detected flu cases in Ingham County is 40\
-someone infected with the flu is infectious for 5 days\
-only 1 in every 3 flu infections is detected.\

3a) What is the analytical probability that no one in our classroom 
has an active flu infection?

\
```{r}
nClass <- 55
ingPop <- 2.8e5
nDaysInf <- 5
avgDailyCases <- 40
detProb <- 1/3
pInf <- (nDaysInf * avgDailyCases / ingPop)/detProb
```
p(a random Ingham County resident has flu) = 
(3 cases for every detected case * 5 days * 40 cases/day) / 280000 people in county = 
`r pInf`
p(a random Ingham resident _does not_ have the flu) = `r 1-pInf`

p(no one in class is infected) = `r (1-pInf)^nClass`

\

3b) What is the probability at least one person in the room **_does_** have an active 
flu infection

p(at least one person is infected) = `r 1-(1-pInf)^nClass`

3c) What is probability 2 people in the room are infected?

`r choose(nClass,2)* (pInf^2) * (1-pInf)^(nClass-2)`

3d) Write a function for simulating this problem for gatherings of arbitrary size.
The arguments of the function should be the number of people in the room 
and the probability that a randomly selected person in the county has an active flu infection,
and it should return the probability (from simulations), 
that no one in the room has an active flu infection.

\
\
```{r}
simFlu <- function(N,pInf,nReps){
	sims <- replicate(nReps,
						any(
							sample(x = c(0,1),
								   size = N,
								   p = c(1-pInf,pInf),
								   replace = TRUE)
							== 1)
					  )
	return(1-length(which(sims))/nReps)
}

simFlu(N=10,pInf=pInf,nReps=1e4)
simFlu(N=nClass,pInf=pInf,nReps=1e4)
simFlu(N=100,pInf=pInf,nReps=1e4)
simFlu(N=500,pInf=pInf,nReps=1e4)
```
\
\