---
title: "Intro to Probability 2: Assignment"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment.Rmd") -->

#### Due by the beginning of next class (10/30)

Save this Rmarkdown script as a file named “YourlastName_Assignment16.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment16.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Define the following terms: 

1a) random variable
 
 **A variable that is the outcome of a random/stochastic process**
 \
 \
 \
 
1b) probability distribution function
 
 **This function describes the probability of outcomes across the sample space**
 \
 \
 \


### Question 2
\
<div class="centered">
$\Large {{n}\choose{k}}p^k (1-p)^{n-k}$
</div>

2a) Using your own words, break down the equation above and 
describe what each part means (or is doing).

**This equation describes the probability distribution function of coin flipping (or similar processes).**
**n is the number of trials (or flips)**\
**k is the number of successes**\
**and p is the probability of success**\

\
**$p^{k} (1-p)^{n-k}$ is the probability of getting k successes and n-k failures given a certain underlying probability k. We get this value through the rule that the probability of a shared event is the product of the probabilities of all the simple events that make up the shared event**
**${n}\choose{k}$ is the number of permutations that will give you k successes. This is the binomial coefficient**
\
\

2b) Using the function of this equation that you coded up in class, calculate the probability of 
seeing 71 successes out of 531 trials, where the probability of success is 0.12345.

```{r, echo=T}

coinFlipProb <- function(n,k,p){
  myprob = choose(n,k)*p^k*(1-p)^(n-k)
  return(myprob)
}

coinFlipProb(n=531,k=71,p=.12345)

```

### Question 3

Data from Dr. Josephs' undergraduate class exams suggests that at any point, 
1% of students at MSU have an active COVID infection.
What is the probability that no one in our classroom would has an active case of COVID?

3a) What is the analytical probability that no one in our classroom 
has an active COVID infection? Assume there are 35 people in the classroom.
\
**First we need to figure out how many people actually have the flu.**\
\
We can repurpose our coin flip function from above. In this case "success" is having COVID and we are interested in the probability of no successes/no flu.\
```{r, echo=T}
pNoFlu = coinFlipProb(n=35, k=0, p=0.01)
pNoFlu
```

3b) What is the probability at least one person in the room **_does_** have an active 
flu infection\
\
**P(Someone has the flu) = 1 - P(no one has the flu)**
```{r, echo=T}
1-pNoFlu
```
\
\
3c) What is probability that 2 people in the room are infected?

**Again we can repurpose the coin flip probability function.

```{r, echo=T}
p2flu = coinFlipProb(35,2,0.01)
p2flu
```



3d) Write a function for simulating this problem for gatherings of arbitrary size.
The arguments of the function should be the number of people in the room 
and the probability that a randomly selected person in the county has an active flu infection,
and it should return the probability (from simulations), 
that no one in the room has an active flu infection.

```{r, echo=T}

## write a function that does one simulation of a class of people who may or may not have the flu
## inputs are prob of flu (myP) and number of people in the class (myN)
sickClass <- function(myP, myN){
  ## sample a class of myN people (1 if have flu, 0 if don't have fl
  myclass = sample(x=c(1,0),size=myN, replace=T, prob = c(myP, 1-myP))
  # return TRUE if someone is sick, if not return FALSE
  mysick = sum(myclass) > 0 
  return(mysick)}
  
##run the sickClass function 1000 times
classSims <- function(myP, myN){
  ## use the replicate function to run the simulations
  allSims = replicate(1e4, sickClass(myP =myP, myN=myN))
  ##  calculate how often someone is sick in all the simulations
  probSick = sum(allSims)/1e4
 return(1-probSick) 
}

classSims(myP = 0.01, myN = 35)

```

A more elegant function for simulating the sickness.
```{r, eval=F, echo=T}
simFlu <- function(N,pInf,nReps){
	sims <- replicate(nReps,
						any(
							sample(x = c(0,1),
								   size = N,
								   p = c(1-pInf,pInf),
								   replace = TRUE)
							== 1)
					  )
	return(1-length(which(sims))/nReps)
}

```