---
title: "Intro to Probability 2: Assignment"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment.Rmd") -->

#### Due by the beginning of next class (10/30)

Save this Rmarkdown script as a file named “YourlastName_Assignment16.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment16.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Re-answer question 2b-e from the previous homework using an analytical approach.


2b) What is the probably that George's birthday is January 1st (Julian date=1)? 

There are 365 possible birthdays

```{r}
1/365
```

2c) Exclusive events: What is the probability that George's birthday is in January (Julian dates 1-31) 

```{r}
31/365
```

2d) Nonexclusive events: What is the probability that George's birthday is in January AND that George was born on a Sunday (assume that there is an equally likely probability of having any day of the week on any date).

```{r}
(31/365)*(1/7)
```

2e) Nonexclusive events: What is the probability that George's birthday is in January OR that George was born on a Sunday?

```{r}
(31/365)+(1/7)-(31/365)*(1/7)
```

## Question 2 -- Probability challenges

You are given an urn containing 100 balls; n of them are red, and 100-n are green, where n is chosen uniformly at random in [0, 100]. You take a random ball out of the urn—it’s red—and discard it. The next ball you pick (out of the 99 remaining) is (from https://x.com/littmath/status/1751648838501224790)

```{r}
#set n
#n = 50

urnSim <- function(){
  
#set n
  n = sample(1:100, size=1)
  
#pick our urn
myUrn = c(rep(1,times=n), rep(0,times=100-n))

#pick two balls
twoBalls = sample(myUrn, size=2, replace=FALSE)


if (twoBalls[1] == 0) {myOut = NA} ## got green first
else if (twoBalls[2] == 0){myOut = 0}  ## got red first and green second
else if (twoBalls[2]==1){myOut=1} ##got red both times

return(myOut)
}

#run it 1000 times
mySims <- replicate(1000, urnSim())

##how often did we get green first?
greenfirst = sum(is.na(mySims))

# how often did we get red, and then green?
sum(mySims==0, na.rm=T)/(1000-greenfirst)

#how often did we get red, and then another red?
sum(mySims==1, na.rm=T)/(1000-greenfirst)

```

Flip a fair coin 100 times—it gives a sequence of heads (H) and tails (T). For each HH in the sequence of flips, Alice gets a point; for each HT, Bob does, so e.g. for the sequence THHHT Alice gets 2 points and Bob gets 1 point. Who is most likely to win? (https://x.com/littmath/status/1769044719034647001)


```{r}
library(stringr)

flipSim <- function(){
  myFlips = sample(c('H','T'), size=100, replace=T)
  myString =  paste0(myFlips, collapse="")
  Alice = str_count(myString, 'HH')
  Bob = str_count(myString,'HT')
  
  if(Alice > Bob){myOut='Alice'}
  if (Bob>Alice){myOut='Bob'}
  if(Bob==Alice){myOut="Tie"}
  
  return(myOut)
  }

mySims <- replicate(1000,flipSim())

sum(mySims=='Bob')/1000

```


### Question 3

Data from Dr. Josephs' undergraduate class exams suggests that at any point, 
1% of students at MSU have an active COVID infection.
What is the probability that no one in our classroom would has an active case of COVID?

3a) What is the analytical probability that no one in our classroom 
has an active COVID infection? Assume there are 35 people in the classroom.
\

The probability of a person having COVID is 1/100, so the probability of that person not having COVID is 99/100.

The probability of two people not having COVID is the probability that the first person doesn't have COVID times the probability that the second person doesn't have COVID, so (99/100)^2

Therefore the probability that all 35 people don't have the flu is (99/100)^35

```{r, echo=T}
(99/100)^35
```


3b) Write a function for simulating this problem for gatherings of arbitrary size.
The arguments of the function should be the number of people in the room 
and the probability that a randomly selected person in the county has COVID ,
and it should return the probability (from simulations), 
that no one in the room has COVID.

```{r, echo=T}

## write a function that does one simulation of a class of people who may or may not have the flu
## inputs are prob of flu (myP) and number of people in the class (myN)
sickClass <- function(myP, myN){
  ## sample a class of myN people (1 if have flu, 0 if don't have fl
  myclass = sample(x=c(1,0),size=myN, replace=T, prob = c(myP, 1-myP))
  # return TRUE if someone is sick, if not return FALSE
  mysick = sum(myclass) > 0 
  return(mysick)}
  
##run the sickClass function 1000 times
classSims <- function(myP, myN){
  ## use the replicate function to run the simulations
  allSims = replicate(1e4, sickClass(myP =myP, myN=myN))
  ##  calculate how often someone is sick in all the simulations
  probSick = sum(allSims)/1e4
 return(1-probSick) 
}

classSims(myP = 0.01, myN = 35)

```

A more elegant function for simulating the sickness.
```{r, eval=F, echo=T}
simFlu <- function(N,pInf,nReps){
	sims <- replicate(nReps,
						any(
							sample(x = c(0,1),
								   size = N,
								   p = c(1-pInf,pInf),
								   replace = TRUE)
							== 1)
					  )
	return(1-length(which(sims))/nReps)
}

```