---
title: "Intro to Bayes: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_20_Assignment_Key.Rmd") -->

#### Due by the beginning of next class (11/16)

Save this Rmarkdown script as a file named “YourlastName_Assignment20.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment20.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about Bayesian statistics

2. extrapolate from in-class exercises to apply Bayesian inference to arbitrary distributions

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

1. Write down Bayes' Theorem using $D$ to denote "data" and $\theta$ to denote "the hypothesis"

$\Large p(\theta \mid D) = \frac{p(D \mid \theta)p(\theta)}{p(D)}$

2. Label the different components using the following list of concepts:
	- the posterior probability of the hypothesis
	- the marginal likelihood of the data
	- the prior probability of the hypothesis
	- the likelihood of the data given the hypothesis

$\Large \text{posterior probability} = \frac{\text{likelihood } \times \text{ prior}}{\text{marginal likelihood}}$

### Question 2

If you know a child younger than 5 (or a parent of a child younger than 5), 
chances are you've heard of RSV (respiratory syncytial virus), 
a flu-like illness that infects virtually all children in their 
first few years of life. Although RSV prevalence usually peaks in the winter 
months in the Northern Hemisphere, this year has seen an unusual 
wave in the late summer/early fall season.

To detect RSV, healthcare providers generally administer antigen tests, 
which are cheaper and faster (although generally less accurate) than PCR-based tests.
The most commonly used RSV antigen test has a 91.7% sensitivity and a 98.9% specificity. 
Let's say my younger kid (Abe) shows some symptoms of RSV, so we get him
an antigen test, which comes back positive. 
Given that result, what is the probability Abe is actually infected with RSV?

In your calculations, assume the following:\

-There are ~7,000 kids under the age of 2 in Ingham County\
-Abe can be treated as a random draw from the <2yos of Ingham County\
-over the last 10 days, the average per-day number of new detected cases of active RSV infection in Ingham County is 5\
-someone infected with RSV is infectious for 7 days\
-only 1 out of every 3 RSV infections is detected and reported.\

prior: $p(\text{rsv})$ = (new daily cases * number of days infectious * number of unreported cases per reported case)/toddlers of Ingham County = `r 3*5*7/7000`\

```{r, echo=FALSE}
priorRSV <- 3*5*7/7000
sensitivity <- 0.917
specificity <- 0.989
falsePos <- 1-specificity
```


likelihood: $p(\text{positive result} \mid \text{infected with RSV})$ = `r sensitivity`\

marginal likelihood: $p(\text{positive result}) = p(\text{rsv})p(\text{positive result} \mid \text{rsv}) + (1-p(\text{rsv}))p(\text{positive result} \mid \text{no rsv}) = 
`r priorRSV*sensitivity + (1-priorRSV)*falsePos`$

posterior probability: $p(\text{rsv} \mid \text{positive result}) = \frac{p(\text{positive result} \mid \text{rsv}) \times p(\text{rsv})}{p(\text{positive result})}$

$\Large p(\text{rsv} \mid \text{positive result}) =  `r (sensitivity * priorRSV)/(priorRSV*sensitivity + (1-priorRSV)*falsePos)`$

So there is only a `r round((sensitivity * priorRSV)/(priorRSV*sensitivity + (1-priorRSV)*falsePos),3)*100`% 
chance that, given his positive antigen test result, Abe is actually infected with RSV.


### Question 3

3a) Building on the results from Question 2, 
write an R function that explores how the probability Abe is actually infected with RSV 
given a positive result from an antigen test 
is dependent on the number of infected toddlers in Ingham County.
The function should take as its arguments: \
 -the percentage of toddlers in Ingham County who are actively infected \
 -the sensitivity of the antigen test (from Question 2)\
 -the specificity of the antigen test (from Question 2)\
 
and it should return the probability a given toddler is infected with RSV given the positive test result.


```{r}

haveRSV <- function(inghamRSV,sensitivity,specificity){
	pHaveRSV <- (sensitivity * inghamRSV) / (inghamRSV * sensitivity + (1-inghamRSV) * (1-specificity))
	return(pHaveRSV)
}

inghamRSV <- seq(0,0.1,length.out=1000)
rsvProbs <- haveRSV(inghamRSV=inghamRSV,
	 		   		  sensitivity=0.917,
	 		   		  specificity=0.989)
plot(inghamRSV,rsvProbs,
	 xlab="background rate of RSV in Ingham County",
	 ylab="probability a given toddler is actually infected RSV\n given a positive test result")

```

3b) At what background level of RSV prevalence in Ingham County does 
the probability of a given toddler actually being infected given a positive antigen test result go above 0.9?

```{r}
inghamRSV[which(rsvProbs>0.9)][1]
```



