---
title: "Skills & Concept Review Homework: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Review_Assignment.Rmd") -->

#### Due by the beginning of next class (11/18)

Save this Rmarkdown script as a file named “YourlastName_ReviewAssignment.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_ReviewAssignment.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability, likelihood, probability distributions, and Bayesian statistics

2. use your R skillz: function building, data visualization, etc.

\

### Question 1

You are at a county fair playing a game of chance. 
The rules are: you roll two 6-sided dice then flip two coins.
If you roll a 7 AND both coins come up heads, you win the prize: 
a large stuffed teddy bear.

1a. Assuming the dice and coins are fair, 
what is the probability that you win the teddy bear?

Answer:
P(roll 7) = number of ways to get 7 / number of possible rolls

number of possible rolls = 6x6 = 36

number of ways to get 7: 1:6, 2:5, 3:4, 4:3, 5:2, 6:1 = 6

P(roll 7) = 1/6

P(2 heads) = P(head) x P(head) = 1/2 x 1/2 = 1/4

P(win) = P(roll 7) x P(2 heads) = 1/6 x 1/4 = 1/24

1b. Write a function that simulates this game, and use it to confirm 
the analytical probability you calculated in part 1a. I've included 
pseudocode to help guide your function-writing.


```{r}

# [twist: the function doesn't require you to pass
#		it any arguments]

carnivalGame <- function(){
  

# first, roll two six-sided dice and add them together
#		(hint: sample)

myDice = sum(sample(1:6, size=2, replace=T))

# next, flip two coins
#		(hint: rbinom)

myCoins = rbinom(n=1, size=2,prob=0.5)

# use a logic statement to determine 
#	whether the outcome is a win or a loss

didIWin = 0 #start with not winning
if (myDice == 7 && myCoins == 2){didIWin <- 1}
return(didIWin)
}

# call the function many times and determine  
#	the probability of winning from the 
#	outcome of the simulations (number of times you won
#	out of total number of simulations)

manySims <- sapply(1:1e5, function(x){carnivalGame()})

probWin = sum(manySims)/1e5
probWin

```

1c. Suppose that the rules of the game have now changed 
and you win the teddy bear if you roll a 7 OR if you flip 
two heads (or if you do both).  Edit the function you wrote in part 1b to 
simulate the probability of winning under this new game.

```{r}


carnivalGame2 <- function(){
myDice = sum(sample(1:6, size=2, replace=T))
myCoins = rbinom(n=1, size=2,prob=0.5)
didIWin = 0 #start with not winning
if (myDice == 7 || myCoins == 2){didIWin <- 1}
return(didIWin)
}

manySims <- sapply(1:1e5, function(x){carnivalGame2()})

probWin = sum(manySims)/1e5
probWin

```


### Question 2

Use maximum likelihood inference to estimate the parameter values 
I used to simulate data from a Normal distribution, 
which are saved as a .Robj file in the Review folder.
Note that you will have to estimate BOTH the mean and the 
standard deviation.

```{r,cache=TRUE}
load("normData.Robj")

# use mean(normData) and sd(normData)
# to inform the parameter ranges within 
# which you do your grid search

mean(normData)
sd(normData)

meanRange <- mean(normData) + c(-abs(mean(normData))/10,abs(mean(normData))/10)
sdRange <- sd(normData) + c(-sd(normData)/10,sd(normData)/10)


# use seq() to define the increments at which you'll 
#	evaluate the likelihood of your data

nSteps = 1e2
myMean = seq(meanRange[1], meanRange[2], length.out = nSteps)
mySD = seq(sdRange[1], sdRange[2], length.out=nSteps)

# create an empty matrix in which to store the log-likelihood
#	of the data for each parameter combination

myMatrix = matrix(data=NA, nrow = nSteps, ncol=nSteps)

# calculate the log likelihood of the data 
#	given every possible parameter combination
for(i in 1:nSteps){
	for(j in 1:nSteps){
		myMatrix[i,j] <- sum(dnorm(x=normData,mean=myMean[i],sd=mySD[j],log=TRUE))
	}
}


# look at it!
image(myMatrix, axes=FALSE)
axis(1, at=seq(0,1,length.out=10),labels = round(seq(meanRange[1], meanRange[2], length.out=10), digits=1))
axis(2, at=seq(0,1,length.out=10),labels = round(seq(sdRange[1], sdRange[2], length.out=10), digits=1))


# report the MLEs
myMean[which(myMatrix==max(myMatrix),arr.ind=TRUE)[1,1]]
mySD[which(myMatrix==max(myMatrix),arr.ind=TRUE)[1,2]]

```

### Question 3

Suppose there are two full bowls of cookies. 
Bowl #1 has 10 chocolate chip and 30 snickerdoodles, 
while bowl #2 has 20 of each. 
Our friend Cookiemonster picks a bowl at random, and then picks a cookie at random. 
We have no reason to believe Cookiemonster treats one bowl differently from another, 
likewise for the cookies. 
The cookie turns out to be a snickerdoodle.

 
3a. Use Bayes Theorem to calculate the analytical probability that 
Cookiemonster picked it out of Bowl #1?


$\normalsize p(bowl1 \mid snickerdoodle) = \frac{p(snickerdoodle \mid bowl  1) \times p(bowl1)}{p(snickerdoodle)}$

$\normalsize p(bowl 1) = 1/2$

$\normalsize p(snickerdoodle \mid bowl) = 3/4$

$\normalsize p(snickerdoodle) = p(bowl 1) \times p(snickerdoodle \mid bowl 1) + p(bowl 2) \times p(snickerdoodle \mid bowl2)$
$\normalsize p(snickerdoodle) = 1/2 \times 3/4 + 1/2 \times 1/2 = 5/8$


$\normalsize p(bowl  1 \mid snickerdoodle) =  \frac{3/4 \times 1/2}{5/8}$

```{r}
((3/4)*(1/2))/(5/8)
```



3b. Write R code to simulate this problem to see 
if your answer from 3a was correct.


```{r}
# define the cookies in bowl 1
bowl1 = c(rep('chocolatechip', 10), rep('snickerdoodle',30))

# define the cookies in bowl 2
bowl2 = c(rep('chocolatechip', 20), rep('snickerdoodle',20))

##I'm gonna stick them in a dataframe to make life easier
bowldf = data.frame(bowl1, bowl2)

# define a function to sample a bowl at random
#	then sample a cookie from that bowl

cookieMonster = function(){
  myN = sample(1:2, size=1)
  myBowl = bowldf[,myN]
  myCookie = sample(myBowl, size=1)
  return(c(myCookie, myN))
}

# simulate that process many times

myCookies = replicate(1e4,cookieMonster())

# think about how to interpret the results!

## pull out all the ones where I got a snickerdoodle
myCookiesdf = data.frame(t(myCookies)) ##easier in a dataframe
library(dplyr)
mySnickerdoodles = dplyr::filter(myCookiesdf, X1 == "snickerdoodle")

## how often is it bowl 1?
myPropBowl1 = nrow(dplyr::filter(mySnickerdoodles, X2 == 1))/nrow(mySnickerdoodles)

myPropBowl1


```

