---
title: "Skills & Concept Review Homework: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Review_Assignment.Rmd") -->

#### Due by the beginning of next class (11/29)

Save this Rmarkdown script as a file named “YourlastName_ReviewAssignment.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_ReviewAssignment.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability, likelihood, probability distributions, and Bayesian statistics

2. use your R skillz: function building, data visualization, etc.

\

### Question 1

You are at a county fair playing a game of chance. 
The rules are: you roll two 6-sided dice then flip two coins.
If you roll a 7 AND both coins come up heads, you win the prize: 
a large stuffed teddy bear.

1a. Assuming the dice and coins are fair, 
what is the probability that you win the teddy bear?

Answer:
P(roll 7) = number of ways to get 7 / number of possible rolls

number of possible rolls = 6x6 = 36

number of ways to get 7: 1:6, 2:5, 3:4, 4:3, 5:2, 6:1 = 6

P(roll 7) = 1/6

P(2 heads) = P(head) x P(head) = 1/2 x 1/2 = 1/4

P(win) = P(roll 7) x P(2 heads) = 1/6 x 1/4 = 1/24

1b. Write a function that simulates this game, and use it to confirm 
the analytical probability you calculated in part 1a. I've included 
pseudocode to help guide your function-writing.


```{r}

# [twist: the function doesn't require you to pass
#		it any arguments]

carnivalGame <- function(){
  

# first, roll two six-sided dice and add them together
#		(hint: sample)

myDice = sum(sample(1:6, size=2, replace=T))

# next, flip two coins
#		(hint: rbinom)

myCoins = rbinom(n=1, size=2,prob=0.5)
## you could also use sum(sample(0:1, size=2, replace=T)) if you want

# use a logic statement to determine 
#	whether the outcome is a win or a loss

didIWin = 0 #start with not winning
if (myDice == 7 && myCoins == 2){didIWin <- 1}
return(didIWin)
}

# call the function many times and determine  
#	the probability of winning from the 
#	outcome of the simulations (number of times you won
#	out of total number of simulations)

manySims <- sapply(1:1e5, function(x){carnivalGame()})

probWin = sum(manySims)/1e5
probWin

```


### Question 2

Use maximum likelihood inference to estimate the parameter values 
I used to simulate data from a Normal distribution, 
which are saved as a .Robj file in the Review folder.
Note that you will have to estimate BOTH the mean and the 
standard deviation.

```{r,cache=TRUE}
load("normData.Robj")

# use mean(normData) and sd(normData)
# to inform the parameter ranges within 
# which you do your grid search

mean(normData)
sd(normData)

meanRange <- mean(normData) + c(-abs(mean(normData))/10,abs(mean(normData))/10)
sdRange <- sd(normData) + c(-sd(normData)/10,sd(normData)/10)


# use seq() to define the increments at which you'll 
#	evaluate the likelihood of your data

nSteps = 1e2
myMean = seq(meanRange[1], meanRange[2], length.out = 1e2)
mySD = seq(sdRange[1], sdRange[2], length.out=1e2)

# create an empty matrix in which to store the log-likelihood
#	of the data for each parameter combination

myMatrix = matrix(data=NA, nrow = nSteps, ncol=nSteps)

# calculate the log likelihood of the data 
#	given every possible parameter combination
for(i in 1:nSteps){
	for(j in 1:nSteps){
		myMatrix[i,j] <- sum(dnorm(x=normData,mean=myMean[i],sd=mySD[j],log=TRUE))
	}
}


# look at it!
image(myMatrix, axes=FALSE)
axis(1, at=seq(0,1,length.out=10),labels = round(seq(meanRange[1], meanRange[2], length.out=10), digits=1))
axis(2, at=seq(0,1,length.out=10),labels = round(seq(sdRange[1], sdRange[2], length.out=10), digits=1))


# report the MLEs
myMean[which(myMatrix==max(myMatrix),arr.ind=TRUE)[1,1]]
mySD[which(myMatrix==max(myMatrix),arr.ind=TRUE)[1,2]]

```

### Question 3

Suppose there are two full bowls of cookies. 
Bowl #1 has 10 chocolate chip and 30 snickerdoodles, 
while bowl #2 has 20 of each. 
Our friend Cookiemonster picks a bowl at random, and then picks a cookie at random. 
We have no reason to believe Cookiemonster treats one bowl differently from another, 
likewise for the cookies. 
The cookie turns out to be a snickerdoodle.

 
3a. Use Bayes Theorem to calculate the analytical probability that 
Cookiemonster picked it out of Bowl #1?


$\normalsize p(bowl1 \mid snickerdoodle) = \frac{p(snickerdoodle \mid bowl  1) \times p(bowl1)}{p(snickerdoodle)}$

$\normalsize p(bowl 1) = 1/2$

$\normalsize p(snickerdoodle \mid bowl) = 3/4$

$\normalsize p(snickerdoodle) = p(bowl 1) \times p(snickerdoodle \mid bowl 1) + p(bowl 2) \times p(snickerdoodle \mid bowl2)$
$\normalsize p(snickerdoodle) = 1/2 \times 3/4 + 1/2 \times 1/2 = 5/8$


$\normalsize p(bowl  1 \mid snickerdoodle) =  \frac{3/4 \times 1/2}{5/8}$

```{r}
((3/4)*(1/2))/(5/8)
```



3b. Bonus! Write R code to simulate this problem to see 
if your answer from 3a was correct.


```{r}
# define the cookies in bowl 1
bowl1 = c(rep('chocolatechip', 10), rep('snickerdoodle',30))

# define the cookies in bowl 2
bowl2 = c(rep('chocolatechip', 20), rep('snickerdoodle',20))

##I'm gonna stick them in a dataframe to make life easier
bowldf = data.frame(bowl1, bowl2)

# define a function to sample a bowl at random
#	then sample a cookie from that bowl

cookieMonster = function(){
  myN = sample(1:2, size=1)
  myBowl = bowldf[,myN]
  myCookie = sample(myBowl, size=1)
  return(c(myCookie, myN))
}

# simulate that process many times

myCookies = replicate(1e4,cookieMonster())

# think about how to interpret the results!

## pull out all the ones where I got a snickerdoodle
myCookiesdf = data.frame(t(myCookies)) ##easier in a dataframe
library(dplyr)
mySnickerdoodles = dplyr::filter(myCookiesdf, X1 == "snickerdoodle")

## how often is it bowl 1?
myPropBowl1 = nrow(dplyr::filter(mySnickerdoodles, X2 == 1))/nrow(mySnickerdoodles)

myPropBowl1


```

Question 4 (Linear Models)

Genome-wide association mapping studies often use a linear model to test for the association between genotype and phenotype.
Specifically, if Y is phenotype and X is genotype, we model Y ~ b*X + a + sigma, where b is the genotypic effect, a is an intercept, and sigma is error.
Genotypes for diploid individuals are coded as 0, 1, or 2.
I've included data for genotypes at 2 loci and a phenotype (fruit count) for a population of Arabidopsis thaliana in the file "gwas-data.rda".


```{r, eval=T, include=F}
##simulate the data
myN=1000
myGenos = sample(0:2, myN, replace=T)
myBeta = 2
myAlpha = 100
mySigma = 10
myPhenos = rnorm(myN, mean=myGenos*myBeta, sd=mySigma) + myAlpha
gwasData = data.frame(genotype = myGenos, phenotype = myPhenos)
myGenos2 = sample(0:2, myN, replace=T)
gwasData = data.frame(locus1 = myGenos, locus2 = myGenos2, phenotype = myPhenos)
save(gwasData, file = 'gwas-data.rda')
```



4a) Load the data and plot a histogram of the phenotypes. Make a scatter plot for each locus that has genotype on the X axis and phenotype on the y axis.
```{r}
#load('gwas-data.rda')

hist(gwasData$phenotype, main="")

plot(gwasData$locus1, gwasData$phenotype)
plot(gwasData$locus2, gwasData$phenotype)

```


4b) Use lm() to separately estimate the effect of genotype at each locus and phenotype. Which locus do you think is more likely to affect phenotype? (hint, you can use summary() of the model output to get more information about the model)

```{r}
myLocus1Model = lm(gwasData$phenotype~gwasData$locus1)
summary(myLocus1Model)

myLocus2Model = lm(gwasData$phenotype~gwasData$locus2)
summary(myLocus2Model)
```

