---
title: "Skills & Concept Review Homework: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Review_Assignment.Rmd") -->

#### Due by the beginning of next class (11/29)

Save this Rmarkdown script as a file named “YourlastName_ReviewAssignment.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_ReviewAssignment.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability, likelihood, probability distributions, and Bayesian statistics

2. use your R skillz: function building, data visualization, etc.

\

### Question 1

You are at a county fair playing a game of chance. 
The rules are: you roll two 6-sided dice then flip two coins.
If you roll a 7 AND both coins come up heads, you win the prize: 
a large stuffed teddy bear.

1a. Assuming the dice and coins are fair, 
what is the probability that you win the teddy bear?



1b. Write a function that simulates this game, and use it to confirm 
the analytical probability you calculated in part 1a. I've included 
pseudocode to help guide your function-writing.



```{r}

# [twist: the function doesn't require you to pass
#		it any arguments]

# first, roll two six-sided dice and add them together
#		(hint: sample)

# next, flip two coins
#		(hint: rbinom)

# use a logic statement to determine 
#	whether the outcome is a win or a loss

# call the function many times and determine  
#	the probability of winning from the 
#	outcome of the simulations (number of times you won
#	out of total number of simulations)

```

1c. Suppose that the rules of the game have now changed 
and you win the teddy bear if you roll a 7 OR if you flip 
two heads (or if you do both).  Edit the function you wrote in part 1b to 
simulate the probability of winning under this new game.


### Question 2

Use maximum likelihood inference to estimate the parameter values 
I used to simulate data from a Normal distribution, 
which are saved as a .Robj file in the Review folder.
Note that you will have to estimate BOTH the mean and the 
standard deviation.

```{r,cache=TRUE}

# use mean(normData) and sd(normData)
# to inform the parameter ranges within 
# which you do your grid search


# use seq() to define the increments at which you'll 
#	evaluate the likelihood of your data

# create an empty matrix in which to store the log-likelihood
#	of the data for each parameter combination


# calculate the log likelihood of the data 
#	given every possible parameter combination

# look at it!


# report the MLEs

```

### Question 3

Suppose there are two full bowls of cookies. 
Bowl #1 has 10 chocolate chip and 30 snickerdoodles, 
while bowl #2 has 20 of each. 
Our friend Cookiemonster picks a bowl at random, and then picks a cookie at random. 
We have no reason to believe Cookiemonster treats one bowl differently from another, 
likewise for the cookies. 
The cookie turns out to be a snickerdoodle.

 
3a. Use Bayes Theorem to calculate the analytical probability that 
Cookiemonster picked it out of Bowl #1?




3b. Write R code to simulate this problem to see 
if your answer from 3a was correct.


```{r}
# define the cookies in bowl 1


# define the cookies in bowl 2

# define a function to sample a bowl at random
#	then sample a cookie from that bowl

# simulate that process many times

# think about how to interpret the results!

```

