---
title: "Skills & Concept Review Homework: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Review_Assignment.Rmd") -->

#### Due by the beginning of next class (11/29)

Save this Rmarkdown script as a file named “YourlastName_ReviewAssignment.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_ReviewAssignment.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability, likelihood, probability distributions, Bayesian statistics, and deterministic functions.

2. use your R skills: function building, data visualization, etc.

\

### Question 1 (Probability)

You are at a county fair playing a game of chance. 
The rules are: you roll two 6-sided dice then flip two coins.
If you roll a 7 AND both coins come up heads, you win the prize: 
a large stuffed teddy bear.

1a. Assuming the dice and coins are fair, 
what is the probability that you win the teddy bear?


1b. Write a function that simulates this game, and use it to confirm 
the analytical probability you calculated in part 1a. I've included 
pseudocode to help guide your function-writing.



```{r}

# [twist: the function doesn't require you to pass
#		it any arguments]

# first, roll two six-sided dice and add them together
#		(hint: sample)

# next, flip two coins
#		(hint: rbinom or sample)

# use a logic statement to determine 
#	whether the outcome is a win or a loss

# return the outcome



# Now, call the function many times and determine  
#	the probability of winning from the 
#	outcome of the simulations (number of times you won
#	out of total number of simulations)

```


### Question 2 (Maximum Likelihood)

I simulated data from a normal distribution and saved is as normData.Robj.

Use maximum likelihood inference with a grid search to estimate the parameter values I used.

Note that you will have to estimate BOTH the mean and the 
standard deviation.

I've given you the code to make the grid already, along with some pseudocode.

```{r}
load("normData.Robj") ##don't forget to use setwd() to go to the directory you saved the file in.

## set ranges for the grids around the sample mean and standard deviation
meanRange <- mean(normData) + c(-abs(mean(normData))/10,abs(mean(normData))/10)
sdRange <- sd(normData) + c(-sd(normData)/10,sd(normData)/10)

## make 100 grid entries within the range for each parameter
nSteps = 1e2
myMean = seq(meanRange[1], meanRange[2], length.out = 1e2)
mySD = seq(sdRange[1], sdRange[2], length.out=1e2)


# create an empty matrix in which to store the log-likelihood
#	of the data for each parameter combination
myMatrix = matrix(data=NA, nrow = nSteps, ncol=nSteps)


# calculate the log likelihood of the data 
#	given every possible parameter combination
# fill these values into myMatrix
# it might help to start by writing a function to calculate the log likelihood for a specific value of mean and sd
# and then use a for loop or apply to use that function on all the inputs in the grid

# look at your log likelihoods across the grid with image()


# report the MLEs

```

### Question 3 (Bayes Theorem)

Suppose there are two full bowls of cookies. 
Bowl #1 has 10 chocolate chip and 30 snickerdoodles, 
while bowl #2 has 20 of each. 
Our friend Cookiemonster picks a bowl at random, and then picks a cookie at random. 
We have no reason to believe Cookiemonster treats one bowl differently from another, 
likewise for the cookies. 
The cookie turns out to be a snickerdoodle.

 
3a. Use Bayes Theorem to calculate the analytical probability that 
Cookiemonster picked it out of Bowl #1?




3b. Write R code to simulate this problem to see 
if your answer from 3a was correct.


```{r}
# define the cookies in bowl 1


# define the cookies in bowl 2

# define a function to sample a bowl at random
#	then sample a cookie from that bowl


myFunction = function(){
  x=sample(c(1,2), size=1, replace=TRUE)
  return(x)
}

myFunction()


sample(c("chocolatechip","snickerdoodle"), size=1, prob=c(0.75,0.25))

# simulate that process many times

# think about how to interpret the results!

```

### Question 4 (Linear Models)

Genome-wide association mapping studies often use a linear model to test for the association between genotype and phenotype.
Specifically, if Y is phenotype and X is genotype, we model Y ~ b*X + a + sigma, where b is the genotypic effect, a is an intercept, and sigma is error.
Genotypes for diploid individuals are coded as 0, 1, or 2.
I've included data for genotypes at 2 loci and a phenotype (fruit count) for a population of Arabidopsis thaliana in the file "gwas-data.rda".

4a) Load the data and plot a histogram of the phenotypes. Make a scatter plot for each locus that has genotype on the X axis and phenotype on the y axis.
```{r}
load('gwas-data.rda')

```


4b) Use lm() to separately estimate the effect of genotype at each locus and phenotype. Which locus do you think is more likely to affect phenotype? (hint, you can use summary() of the model output to get more information about the model)
