---
title: "Skills & Concept Review Homework: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Review_Assignment_Key.Rmd") -->

#### Due by the beginning of next class (11/16)

Save this Rmarkdown script as a file named “YourlastName_ReviewAssignment.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_ReviewAssignment.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability, likelihood, probability distributions, and Bayesian statistics

2. use your R skillz: function building, data visualization, etc.

\

### Question 1

You are at a county fair playing a game of chance. 
The rules are: you roll two 6-sided dice then flip two coins.
If you roll a 7 AND both coins come up heads, you win the prize: 
a large stuffed teddy bear.

1a. Assuming the dice and coins are fair, 
what is the probability that you win the teddy bear?

$p(\text{sum(dice)=7 AND two heads}) = p(\text{sum(dice)=7}) \times p(\text{two heads})$

$p(\text{winning}) = \frac{1}{6} \times \frac{1}{4} = `r round(1/24,3)`$

1b. Write a function that simulates this game, and use it to confirm 
the analytical probability you calculated in part 1a. I've included 
pseudocode to help guide your function-writing.



```{r}

# [twist: the function doesn't require you to pass
#		it any arguments]

# first, roll two six-sided dice and add them together
#		(hint: sample)

# next, flip two coins
#		(hint: rbinom)

# use a logic statement to determine 
#	whether the outcome is a win or a loss

# call the function many times and determine  
#	the probability of winning from the 
#	outcome of the simulations (number of times you won
#	out of total number of simulations)

simGame <- function(){
	rolls <- sum(sample(1:6,2,replace=TRUE))
	flips <- rbinom(1,2,prob=0.5)
	if(rolls == 7 && flips==2){
		outcome <- 1
	} else {
		outcome <- 0
	}
	return(outcome)
}

sims <- replicate(1e5,simGame())
length(which(sims==1))/length(sims)
```

1c. Suppose that the rules of the game have now changed 
and you win the teddy bear if you roll a 7 OR if you flip 
two heads (or if you do both).  Edit the function you wrote in part 1b to 
simulate the probability of winning under this new game.

```{r}
simGame <- function(){
	rolls <- sum(sample(1:6,2,replace=TRUE))
	flips <- rbinom(1,2,prob=0.5)
	if(rolls == 7 | flips==2){
		outcome <- 1
	} else {
		outcome <- 0
	}
	return(outcome)
}

sims <- replicate(1e5,simGame())
length(which(sims==1))/length(sims)
```


### Question 2

Use maximum likelihood inference to estimate the parameter values 
I used to simulate data from a Normal distribution, 
which are saved as a .Robj file in the Review folder.
Note that you will have to estimate BOTH the mean and the 
standard deviation.

```{r,echo=FALSE,}
normData <- rnorm(1e4,mean=-16,sd=12)
save(normData,file="normData.Robj")
```


```{r,cache=TRUE}
load("normData.Robj")

# use mean(normData) and sd(normData)
# to inform the parameter ranges within 
# which you do your grid search

meanRange <- mean(normData) + c(-abs(mean(normData))/10,abs(mean(normData))/10)
sdRange <- sd(normData) + c(-sd(normData)/10,sd(normData)/10)

# use seq() to define the increments at which you'll 
#	evaluate the likelihood of your data
nSteps <- 1e2
meanSeq <- seq(meanRange[1],meanRange[2],length.out=nSteps)
sdSeq <- seq(sdRange[1],sdRange[2],length.out=nSteps)

# create an empty matrix in which to store the log-likelihood
#	of the data for each parameter combination

logLik <- matrix(NA,nrow=nSteps,ncol=nSteps)

# calculate the log likelihood of the data 
#	given every possible parameter combination
for(i in 1:nSteps){
	for(j in 1:nSteps){
		logLik[i,j] <- sum(dnorm(x=normData,mean=meanSeq[i],sd=sdSeq[j],log=TRUE))
	}
}

# look at it!
image(logLik)

# report the MLEs

meanSeq[which(logLik==max(logLik),arr.ind=TRUE)[1,1]]
sdSeq[which(logLik==max(logLik),arr.ind=TRUE)[1,2]]
```

### Question 3

Suppose there are two full bowls of cookies. 
Bowl #1 has 10 chocolate chip and 30 snickerdoodles, 
while bowl #2 has 20 of each. 
Our friend Cookiemonster picks a bowl at random, and then picks a cookie at random. 
We have no reason to believe Cookiemonster treats one bowl differently from another, 
likewise for the cookies. 
The cookie turns out to be a snickerdoodle.

 
3a. Use Bayes Theorem to calculate the analytical probability that 
Cookiemonster picked it out of Bowl #1?

$\text{p(bowl1 | snickerdoodle)} = \frac{\text{p(snickerdoodle | bowl1)} \times \text{p(bowl1)}}{\text{p(snickerdoodle)}}$

$\text{p(snickerdoodle | bowl1)} = \frac{30}{40} = `r 30/40`$

$\text{p(bowl1)} = \frac{1}{2} = `r 1/2`$

$\text{p(snickerdoodle) = p(bowl1)} \times \text{p(snickerdoodle | bowl1) + p(bowl2)} \times \text{p(snickerdoodle | bowl2)}$

$\text{p(snickerdoodle)} = \frac{1}{2} \times \frac{30}{40} + \frac{1}{2} \times \frac{20}{40} = `r 1/2 * 30/40 + 1/2 * 20/40`$

$\text{p(bowl1 | snickerdoodle)} = `r (30/40 * 1/2) / (1/2 * 30/40 + 1/2 * 20/40)`$



3b. Write R code to simulate this problem to see 
if your answer from 3a was correct.


```{r}
# define the cookies in bowl 1
bowl1 <- c(rep("chocchip",10),rep("snickerdoodle",30))

# define the cookies in bowl 2
bowl2 <- c(rep("chocchip",20),rep("snickerdoodle",20))

# define a function to sample a bowl at random
#	then sample a cookie from that bowl
simCookie <- function(bowl1,bowl2){
	bowl <- sample(c("bowl1","bowl2"),1)
	if(bowl=="bowl1"){
		cookie <- sample(bowl1,1)
	} else if (bowl=="bowl2"){
		cookie <- sample(bowl2,1)
	}
	return(c(bowl,cookie))
}

# simulate that process many times
nSims <- 1e4
sims <- replicate(nSims,simCookie(bowl1,bowl2))

# count the number of simulations in which Cookiemonster
#	draws a snickerdoodle from bowl1, divided by the number 
# 	of times Cookiemonster draws a snickerdoodle

length(which(sims[1,which(sims[2,]=="snickerdoodle")]=="bowl1"))/length(which(sims[2,]=="snickerdoodle"))

```

