---
title: "Intro to Probability 1: Assignment"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_15_Assignment_key.Rmd") -->

#### Due by the beginning of next class (10/28)

Save this Rmarkdown script as a file named “YourlastName_Assignment15.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment15.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1

Define the following terms: 

 - probability
 
 \
 **How often we expect an event to happen**
 \
 
 - simple event 
 
 \
 ** a single experiment (drawing a random number) with a single outcome (the number drawn)**
 \
 
 - complex event
 
 \
 **a composite of simple events**
 \
 
 - shared event
 
 \
 **the simultaneous occurrence of simple events**
 \
 
 - conditional probability

 \
** the probability of one outcome _conditional_ on another**
 \
 
 - statistical independence

 \
**events are independent when their outcomes do not affect each other**
 \
 \

### Question 2

2a) Suppose a process has two possible outcomes: A and B.
If the probability of A ($p(A)$) is 0.2, what is p(B)?

\
0.8
\

2b) More generally, if a process has many possible outcomes, 
one of which is A, what is the probability of _not_ A?

\
$1-p(A)$
\
\


### Question 3

**The Birthday Problem** is a classic problem in probability:

how many people must be in a room for there to be > 50% probability 
that _at least_ two of them share a birthday?

\

3a) Write a function for simulating the birthday problem. 
The argument of the function should be the number of people in the room, 
and it should return the probability (from simulations), that two or more 
share a birthday.

\
\
```{r,echo=TRUE}

sim.bday <- function(N){
	prop.twinsies <- sum(
					replicate(1e5,
						length(
							unique(
								sample(1:365,N,replace=TRUE)
							)
						) < N)
				   )/1e5
	return(prop.twinsies)
}

```
\
\

3b) Using this function, determine how many people must be in a room for 
there to be > 50% probability that _at least_ two of them share a birthday

\
\
```{r,echo=TRUE}
which(sapply(1:30,function(n){sim.bday(n)}) > 0.5)[1]
```
\
\

3c) Suppose humans colonize Mars.  How many people have to be in a room 
on Mars for there to be a >50% probability that at least 2 of them share 
a **Martian** birthday?

\
\
```{r,cache=TRUE,echo=TRUE}
sim.mars.bday <- function(N){
	prop.twinsies <- sum(
					replicate(1e5,
						length(
							unique(
								sample(1:687,N,replace=TRUE)
							)
						) < N)
				   )/1e5
	return(prop.twinsies)
}

which(sapply(1:33,function(n){sim.mars.bday(n)}) > 0.5)[1]
```
\
\
