---
title: "Intro to Probability 1: Assignment"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_15_Assignment.Rmd") -->

#### Due by midnight on Oct 23 (the night before class on Oct 24)

Save this Rmarkdown script as a file named “YourlastName_Assignment15.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment15.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about probability

2. use simulations to build intuition

3. use your R skillz: function building, data visualization, etc.

\

### Question 1



Write an example of the following statistical concepts:

 - mutually-exclusive events
 
Birthdays -- my birthday cannot be both April 17 and April 20.
 
 -- non-exclusive events
 
Birthdays -- my birthday can be April 17 and I could have been born on a Thursday.
 
 -- conditional probability
 
The probability that my birthday is 24 given that my birthday falls on a Thursday this year. 
 

### Question 2

2a) Write a function simulating sampling a number (N) of dates in Julian date (dates are numbers from 1 to 365) with replacement.

```{r, echo=T}
dateFunction <- function(myN){
  myDates <- sample(1:365, myN, replace=T)
  return(myDates)
}
```

2b) Assume that the probability of being born is the same every day of the year. You make a new friend named George. What is the probably that George's birthday is January 1st (Julian date=1)? Use the function you wrote above to sample this person being born 1e5 times to estimate the probability.

```{r, echo=T}
myBirthdays = replicate(1e5,dateFunction(1))
sum(myBirthdays == 1)/length(myBirthdays)

```

2c) Exclusive events: What is the probability that George's birthday is in January (Julian dates 1-31) Again, do 1e5 simulations to estimate the probability
```{r, echo=T}
sum(myBirthdays < 32)/length(myBirthdays)
```


2d) Nonexclusive events: What is the probability that George's birthday is in January AND that George was born on a Sunday (assume that there is an equally likely probability of having any day of the week on any date).
```{r}
dayFunction <- function(myN){
  myDays <- sample(1:7, myN, replace=T)
  return(myDays)
}

#sample the days of the week that go with every birthday
myDays = replicate(1e5, dayFunction(1))

isJanuary <- myBirthdays < 32
isSunday <- myDays == 1

isJanuaryandSunday = isJanuary + isSunday

sum(isJanuaryandSunday ==2)/1e5


```

### Question 3 (challenge question!!!)

**The Birthday Problem** is a classic problem in probability:

how many people must be in a room for there to be > 50% probability 
that _at least_ two of them share a birthday?

\

3a) Write a function for simulating the birthday problem. 
The argument of the function should be the number of people in the room, 
and it should return the probability (from simulations), that two or more 
share a birthday.


This is how I solved the problem:
```{r, echo=T}

## Write a function where you
birthdaySharing <- function(myN){
  #sample n birthdays from the year (with replacement!)
  birthdays = dateFunction(myN)
  # if anyone shares a birthday, log it as a "birthday twins" event
  if (sum(duplicated(birthdays))>0) { myOut = TRUE} else {myOut = FALSE}
  return(myOut)
}

## run the function 1e4 times
birthdaySim <- function(myP){ ##myP is the number of individuals
myOut = sapply(1:1e4,function(x){birthdaySharing(myP)})##
  return(sum(myOut)/1e4)}

birthdaySim(10)
```

Here is another, more elegant but hard-to-read solution:

```{r,echo=TRUE}
sim.bday <- function(N){
	prop.twinsies <- sum(
					replicate(1e4,
						length(
							unique(
								sample(1:365,N,replace=TRUE)
							)
						) < N)
				   )/1e5
	return(prop.twinsies)
}
```

```

3b) Using this function, determine how many people must be in a room for 
there to be > 50% probability that _at least_ two of them share a birthday

```{r, echo=T}

which(sapply(1:30,function(n){birthdaySim(n)}) > 0.5)[1]

```


3c) Suppose humans colonize Mars.  How many people have to be in a room 
on Mars for there to be a >50% probability that at least 2 of them share 
a **Martian** birthday?

\
