---
title: "Intro to Probability 2: Assignment"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_16_Assignment.Rmd") -->

#### Due by midnight the day before the next class (10/30)

Save this Rmarkdown script as a file named “YourlastName_Assignment17.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment17.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

1) We'll revisit a question from the previous homework but use our binomial distribution skills to answer it. 
This time, instead of calculating the probability directly, use the coin flip function from
today's class to calculate the probability.
\
\

a) Data from Emily's undergraduate class exams suggests that at any point, 
1% of students at MSU have an active COVID infection.
What is the probability that no one in our classroom (n=25) has an active case of COVID? 

\

```{r}
coinFlipProb <- function(n,k,p){
  myprob = choose(n,k)*p^k*(1-p)^(n-k)
  return(myprob)
}

pNoCovid = coinFlipProb(n=25, k=0, p=0.01)

pNoCovid
```

b) What is the probability that one person in the room has an active COVID infection?

```{r}

pOneCovid = coinFlipProb(n=25, k=1, p=0.01)

pOneCovid
```

c) Make a plot of the probability distribution with number of infected people on the x axis and probability on the y axis

```{r}


plot(0:25, coinFlipProb(n=25, k=0:25, p=0.01),
     xlab = "number of infections", ylab = "probability")


```


2) R has built in functions for the probability distributions. Take a second to exam the dbinom function.

a) Make the same plot from 1c using dbinom instead of the coin flip function.

```{r}

plot(0:25, dbinom(0:25, size=25, prob=0.01),
     xlab = "number of infections", ylab = "probability")

```


