---
title: "Simulation_Battle_Assignment"
output: html_document
date: "2023-11-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Group Name:** DodoCode 

#### Members: Max Halliday, Saniya Henderson, Jenna Hynes, Preethi Muruganandam, and Abigail Palmisano

#### Probability Distribution = Exponential Deterministic Function = Linear Link Function = Exponential (-z\^-1)

#### Explanation of dataset: Dodo birds were discovered in the year 1598, over the course of \~1000 years after their discovery the population quickly declined due to people hunting them for sport. Our dataset was created to show the increase in Dodo bird population before human discovery.

```{r}


set.seed(2) #Determing the range for the predictor variables in an exponential distribution
ran <- (range(500,1598)) #year range before Dodo bird discovery in 1598
data <- rexp(1000,0.01) 
sdata <- data[order(-data)]
timeline <- (500 - 1598) * (sdata - min(sdata)) / (max(sdata) - min(sdata)) + 1598 

detFunc <- function(x,a,b){ #defining a linear deterministic function
  return(a + b*x)
}

neg.inv.link <- function(p){ #specifying inverse link function for an exponential distribution
  return(-p ** -1)
}


dodo_pop <- rexp(neg.inv.link(detFunc(timeline,1e12,1e13))) #simulating response variable (dodo bird population (in thousands))

plot(dodo_pop ~ timeline, xlab = "Time (Years)", ylab = "Population (in the thousands)", 
     main = "Dodo Bird Population Before Human Discovery", col = "seagreen")

myData <- data.frame("predictor" = timeline, "response" = dodo_pop)


```
