---
title: "R Bug Team"
author: "William Smith, Bernard Frantz, Jacob Fitzharris, Ignatius Andika"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Deathoppers
Planthoppers (Cicadellidae) are a diverse family of Hemipterans known to vector many plant diseases. Pesticides are a common control method but pesticides do not kill all individuals in a population. Our data describes persistence of pesticides within the hemolymph of surviving planthoppers after a spray. Our x-axis is the time after spray in minutes our y-axis is the amount of pesticide within the hemolymph.
```{r, echo = T}
set.seed(123)

N0 <- 6
decay <- 0.18

x <- rnorm(1000)

# Inverse function for the exponential decay
inv.logit.link <- function(p){
	x <- exp(p)
}

response <- rpois(1000, lambda = inv.logit.link(N0 * exp(-decay*x)))
predictor <- rnorm(1000, mean = 5.2, sd = 0.65)

data_frame <- as.data.frame(cbind(response, predictor))
plot(data_frame$response~data_frame$predictor,xlab="Time After Spray in Minutes",ylab="Amount of Pesticide in Hemolymph")

write.csv(data_frame, file = "rbug-data.csv")
```

