---
title: "Highly contagious ancient pathogens in permafrost"
subtitle: "IBIO830 Group Project - Simulation/Inference Battle"
author: 
- The R Meltdowns (Team 5) - Jack Magee, Bethany Mikles, Michele Remer, Michelle Volk, Kelly Waters, Esther Wong  
output:
  html_document: 
    df_print: paged
date: "2022-12-01"
---

- Stats-inspired team name: __The R Meltdowns__\
- Probability distribution: __Normal distribution__\
- Deterministic function: __Exponential decay__\
- Link function: Identity $X\beta = \mu$\

Biological scenario:

After an unusually warm fall, citizens of Poissontown, Siberia (population 1000) are suddenly developing disease symptoms from an unknown origin. Laboratory tests have come back negative for influenza, COVID-19, measles, and other known infectious diseases. Scientists descend upon the town and discover a 48,500-year-old virus* in patient respiratory fluids. It appears that increasing temperatures have resulted in the melting of nearby permafrost, releasing ancient pathogens that human immune systems have never been exposed to. The virus, named Bayesviridae, seems to spread via respiratory droplets and is highly contagious. Epidemiologists use a standard SIR model (S=Susceptible, I=Infected, R=Recovered) to estimate how quickly the virus spreads, how long individuals are infectious, and other epidemiological parameters such as the R0, or reproductive number. Our dataset represents the S (susceptible) part of the SIR model. It consists of the number of susceptible individuals recorded from Poissontown during a period of 42 days since the start of the outbreak. As the virus spreads through the population, we expect the number of susceptible individuals to decrease exponentially.\

*Alempic et al 2022 just published their finding that viruses remain infectious after 48,500 years spent in permafrost!

\

```{r}
set.seed(11)
# define a deterministic function - exponential decay
myDet <- function(x,a,b){
    return(a * (x^-b))
}

# define an inverse link function, which should be Identity because our data are 
invLink <- function(z){
   return(z)
}

# define parameters
n = 1000
x = 1:n
a = 1
b = 0.2
s = 0.05
y = as.integer(rnorm(n, mean = invLink(myDet(x, a, b)), sd = s) * 1000)

plot(x = x,
     y = y,
     pch = 19,
     ylab = "Number of Susceptible Individuals", 
     xlab = "Time (hours)")
```


```{r}
# make our dataset
mydata = data.frame("predictor" = x, "response" = y)
mydata
```

```{r}
# export our dataset
write.csv(mydata, "IBIO830_team5_data.csv", row.names = F)
```