---
title: "Simulation/Inference Battle Assignment!"
output:
  html_document:
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->

<!-- rmarkdown::render("SI_battle_assignment.Rmd") -->

```{r,echo=FALSE}
	#set any global options
	options(digits=3)
	set.seed(123)
```

## Frogs. Where are they and in what amount?

The National Science Foundation, being as invested in questions of basic
science as they are, unilaterally funds research about frogs and where
they're located. Our work (as described in our generously awarded NSF
proposal) requires that we count the number of frogs we observe as the
distance between ourselves and the pond shore increases. Ultimately two
things were recorded: **no. of frogs** and **distance from pond** (in
meters).

Incredibly, and defying all known laws of nature, we observed that the
number of total frogs observed **decreased** as we traveled further and
further away (0-100m) from the pond.

Data collection was limited by worsening black bear attacks. We lost one
(1) team member.

```{r echo=FALSE}
n.samples <- 1000
predictor <- runif(n.samples, min = 0, max = 100)
alpha <- 4.5
beta <- -0.03
response <- rpois(n.samples,lambda=exp(alpha*predictor^beta))
frogs <- data.frame(predictor,response)

library(ggplot2)
library(dplyr)

frogs %>% ggplot() + 
  geom_point(aes(x = predictor, 
                 y = response), 
             color = 'black') +
  theme_bw() +
  labs(x = 'Distance in meters (Predictor)',
       y = 'Frogs counted (Response)',
       title = 'Extremely Important Frog Thing')
```

```{r echo=FALSE}
frogs %>% ggplot() + 
  geom_histogram(aes(x = response), 
                 color = 'black',
                 binwidth = 1) +
  theme_bw() +
  labs(x = 'Frogs Counted (Response)',
       y = 'Frequency',
       title = 'Extremely Important Frog Thing')
```

```{r echo=FALSE}
write.csv(frogs, "frogs.csv")
```
