---
title: "Lecture 18 - The World of Probability Distributions"
author: "Emily Josephs"
date: "October 31 2024"
output:
  ioslides_presentation:
    transition: 0.001
    bigger: true
    incremental: true
---
<!-- To render the lecture in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("lecture17.Rmd") -->

```{r,echo=FALSE}
	#set any global options
	options(digits=3)
	set.seed(123)
```


```{r, eval=F, echo=F}
flips <- function(nFlips){
	flips <- sample(c('H','T'), nFlips, replace=TRUE)
	return(flips)
}

flips(3)

#mySims = replicate(100, flips(3), simplify=TRUE)
```

## Today: Probability distributions, continued!

What if the probability of different outcomes varies?

How do we describe this situation?

How can we describe a variety of different types of processes using probability distributions?

## Recap from last time

1. The outcome of a random process is called a **_random variable_**

2. To calculate the probability of any given value of the random variable:
	+ we can either define the state space and count or
	+ we can figure out the probability distribution function

3. A **_probability distribution function_** is a rule or formula 
that describes the probability of all possible outcomes of a random event

4. The **_binomial distribution_** describes the probabilities of success in a set of trials.


## The Binomial Distribution {.build}

Given:

 - a number of trials (**_n_**)
 
 - and a probability of success (**_p_**)

can we come up with a rule for calculating 
the probability of any number of successes (**_k_**)?

\

E.g., what's the probability of 43 successes out of 100 trials?

Worksheet time!


## The Binomial distribution  {.build}

Well, we've just derived the **_probability distribution function_**.

<div class="centered">
$\Large {{n}\choose{k}} \times \left( p^k \times (1-p)^{n-k}	\right)$
</div>

## The Binomial distribution  {.build}

Why are we multiplying the two terms together?

Imagine that we have 2 flips and we want to know how likely we are to get 1 heads?

Our state space = $[H,H], [H,T], [T,H], [T,T]$

All of these options are equiprobable!

$P(2~\text{heads}) = P([H,T]) + P([T,H])$

$P(2~\text{heads}) = 2 \times P([H,T])$

```{r}
choose(n=2, k=1)
```


## The Binomial distribution  {.build}


```{r, echo=T}
	coin.flip.prob <- function(n,k,p){
		prob <- choose(n,k) * p^k * (1-p)^(n-k)
		return(prob)
	}
```
	
```{r, echo=T}
	coin.flip.prob(n=100,k=43,p=0.5)
```

## Coin-flipping {.build}
```{r,fig.width=6.5,fig.height=5,fig.align="center",echo=T}
	plot(coin.flip.prob(n=100,k=0:100,p=0.5),
			xlab="successes",ylab="probability")
```

## Coin-flipping {.build}
```{r,fig.width=6.5,fig.height=5,fig.align="center",echo=T}
	plot(coin.flip.prob(n=100,k=0:100,p=0.3),
			xlab="successes",ylab="probability")
```


## 
```{r,fig.width=6.5,fig.height=5,fig.align="center",echo=FALSE}
	plot(0,type='n',xlim=c(0,25),ylim=c(0,0.2),
			xlab="successes",ylab="probability",main="")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.3),lwd=3,col="blue")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.5),lwd=3,col="purple")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.7),lwd=3,col="red")
		legend(x="topright",lty=1,lwd=3,col=c("blue","purple","red"),legend=c(0.3,0.5,0.7),title="p(heads) = ")
```


## Probability distributions: the binomial {.build}

If a random variable **_X_** is binomially distributed, we write:

\

<div class="centered">
$\Large X \sim B(n,p)$
</div> 


Read as: "_X_ is a draw from a binomial distribution 
with parameters _n_ and _p_"


And, if **_X_** is a draw from a binomial distribution, 
we can use the **_binomial probability distribution function_** 
to calculate the probability for any value of **_X_ = k**, given **_n_** and **_p_**.

## Probability distributions: the binomial {.build}

There are also tools available for working with the binomial distribution:

```{r,eval=FALSE}
# look at rbinom and dbinom
?rbinom
```

Using `rbinom`, generate 1 sample of 100 draws from a 
binomial distribution with `p=0.6`.

```{r}
rbinom(n=1,size=100,p=0.6)
rbinom(n=1,size=100,p=0.6)
```

## Team exercise
Use `rbinom` to generate 1000 samples of 100 draws from the binomial distribution. Pick your own value of *p*.

Plot the histogram of the values that you get.

Calculate the mean of the values.

Calculate the variance of the values (hint: use `var()`).

## Probability distributions: the binomial {.build}

The mean and variance of the binomial distribution are 
defined!

If $X \sim B(n,p)$, we know:

<div class="centered">
$\Large \mathbb{E}[X] = np$
</div> 

and

<div class="centered">
$\Large \mathbb{V}ar(X) = np(1-p)$
</div>

Check to make sure that your estimates from the teams match these values.

## Wait what's a variance? {.build}

The variance tells you about the amount of spread in the data.

$\large \mathbb{V}ar(X) = \mathbb{E}(x_{i}-\bar{x})^{2}$

The variance will be higher if things tend to be farther away from the mean.

(hint: Wikipedia is your friend for math concepts like variance)

## Probability distributions {.build}

The binomial is only one probability distribution; 
there are many others!

Why are they useful?

Because they describe processes that shape 
the world around us!

## Distributions and Random Variables {.build}

Random variables come in two flavors: **_discrete_** and **_continuous_**

- **Discrete** random variables are things that can be counted
	+ heads in a coinflip experiment, fish in a lake, flowers on a plant
\
\

- **Continuous** random variables are things that can be measured
	+ height or weight

\
\

## Distributions and Random Variables {.build}

Discrete probability distributions have 
**_probability mass functions_**

- <div class="centered">
$\Large {{n}\choose{k}}p^k (1-p)^{n-k}$
</div>

- a **_pmf_** gives the probability that a discrete random variable is _exactly_ equal to some value
	+ e.g., what's the probability of getting _exactly_ 43 heads out of 100 coin flips
\

What about continuous RVs?

## {.columns-2 .build}
### Discrete (heads):

\
\

```{r,fig.width=4,fig.height=4,echo=FALSE,align="center"}
plot(coin.flip.prob(n=100,k=0:100,p=0.5),
			xlab="successes",ylab="probability",
			pch=20,col=adjustcolor(1,0.7),main="discrete")
```

\
\

### Continuous (height):

\
\

```{r,fig.width=4,fig.height=4,echo=FALSE,align="center"}
norm.prob <- function(x,mean,sigma){
 	prob <- exp(-((x-mean)^2/(2*sigma^2)))/sqrt(2*pi*sigma^2)
 	return(prob)
}
lower <- 10
upper <- 130
plot(lower:upper,norm.prob(lower:upper,mean=70,sigma=20),type='n',
		xlab="height",ylab="probability",main="continuous")
	lines(seq(lower,upper,length.out=500),
			norm.prob(seq(lower,upper,length.out=500),mean=70,sigma=20))
```

## Probability density functions {.build}

A continuous variable can take 
an infinite possible number of values, 
**the probability of any specific value is 0**.
 
So, a pdf measures the _relative_ probability 
of a given value.

<div class="centered">
```{r,fig.width=5,fig.height=4,echo=FALSE,align="center"}
norm.prob <- function(x,mean,sigma){
 	prob <- exp(-((x-mean)^2/(2*sigma^2)))/sqrt(2*pi*sigma^2)
 	return(prob)
}
lower <- 10
upper <- 130
plot(lower:upper,norm.prob(lower:upper,mean=70,sigma=20),type='n',
		xlab="height",ylab="probability",main="continuous")
	lines(seq(lower,upper,length.out=500),
			norm.prob(seq(lower,upper,length.out=500),mean=70,sigma=20))
```
</div>

## Properties of probability distributions {.build}

+ Mean

+ Variance

+ Discrete vs continuous

+ Boundaries

## Probability distribution boundaries {.build}

What are potential values of the binomial distribution?

```{r,fig.width=6.5,fig.height=5,fig.align="center",echo=FALSE}
	plot(0,type='n',xlim=c(0,25),ylim=c(0,0.2),
			xlab="successes",ylab="probability",main="")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.3),lwd=3,col="blue")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.5),lwd=3,col="purple")
		lines(0:25,coin.flip.prob(n=25,k=0:25,p=0.7),lwd=3,col="red")
		legend(x="topright",lty=1,lwd=3,col=c("blue","purple","red"),legend=c(0.3,0.5,0.7),title="p(heads) = ")
```

## Properties of probability distributions

+ Mean

+ Variance

+ Discrete vs continuous

+ Boundaries

+ Parameters

## Distribution parameters {.build}

What are the parameters of the binomial distribution?

$X \sim B(n,p)$

## Distribution parameters {.build}

What are the parameters of the normal distribution?

$X \sim N(\mu,\sigma)$


## Intro to random variables: recap

1. The outcome of a random process is called a **_random variable_**

2. To calculate the probability of any given value of the random variable:
	+ we can either define the state space and count or
	+ we can figure out the probability distribution function

3. A **_probability distribution function_** is a rule or formula 
that describes the probability of all possible outcomes of a random event

## Intro to probability distributions: recap 

1. There are many named **_probability distributions_**. These are useful 
for describing different natural processes.

2. There are many tools for working with different probability distributions, 
including tools for simulating data, and calculations for the expected mean and variance

3. Probability distributions can be either discrete or continuous.
	+ the probability of discrete events can be characterized by a **_probability mass function_**
	+ the probability of any value in a continuous distribution is 0, but the _relative_ probability 
	can be described by a **_probability density function_**


## Distribution map {.build}

Work in a team of 2-4 to make a guide/decision tree/map/Cosmo quiz to choose the distribution that best matches a dataset.

Here are the distributions your map should help you choose between.

**continuous**: beta, exponential, gamma, lognormal, normal, uniform

**discrete**: poisson, geometric, negative binomial

Use the properties of distributions that we have already talked about: discrete/continuous, mean, variance, boundaries.



