---
title: "Model Building: putting it all together"
output:
  html_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_24_Assignment.Rmd") -->

#### Due by 11:59 on Nov 21.

Save this Rmarkdown script as a file named “YourlastName_Assignment21.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment21.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

Question 1:

Going back to our mite example, when fungus outbreaks are very severe, they can kill the entire plant. 

You are interested in quantifying the relationship between fungus amount and plant mortality in vineyards. 
You collect data on the average amount of fungus per plot and, at the end of the season, count the number of dead plants.
Assume all plots are the same size.

The data is available on d2l to download. 

```{r}
mydata <- read.csv('fungusMortality.csv')
```

1a) Make a plot showing the amount of fungus on the X axis and the number of dead plants on the Y axis. 



1b) Explain why a Poisson distribution would be an appropriate choice for modelling this data.




1c) Head to wikipedia and look up the link function for a Poisson regression: https://en.wikipedia.org/wiki/Generalized_linear_model#Link_function

Below is code for running a linear normal regression. Modify it to run a Poisson regression:

```{r}

#define a deterministic funciton
detFunc <- function(x,a,b){
    return(a + b*x)
}

# specify the inverse link function
invLink <- function(z){
    return(z)
}

# run a regression with mle2()
mod <- mle2(mortality ~ dnorm(mean=invLink( detFunc(dry.wt.fungus,a,b)), sd=s),
            data=mydata,
            start=list("a"=0,"b"=0, 's'=1))

## use the model to generate data
modData = invLink(mod@coef[1]+mydata$dry.wt.fungus*mod@coef[2])

## look at the model-predicted data curve on top of the actual data
plot(mydata$dry.wt.fungus, mydata$mortality, bty="n", col="darkgray")
points(mydata$dry.wt.fungus, modData,col="red",lwd=2)

```

Calculate the residuals and make a plot of the residuals against the independent variable.

Do you think the model is doing a good job of fitting the data? Why or why not?

