---
title: "Model Building: putting it all together"
output:
  html_document: default
  pdf_document: default
language: R
---

<!-- To render the assignment in Rmarkdown, enter the command below in the R console -->
<!-- rmarkdown::render("Class_24_Assignment_key.Rmd") -->

#### Due by the beginning of next class (12/3)

Save this Rmarkdown script as a file named “YourlastName_Assignment24.Rmd”. 
For each question, fill out the answer and, where requested, 
provide the relevant R code using "```" and the echo = TRUE argument.

When finished, Knit your script together into an html report, 
saved as “YourlastName_Assignment24.html” and upload the 
resulting file to the D2L site (in today’s class folder) to turn it in. 
It is due before our next class.

\

**This homework assignment will test your abilities to:**

1. think critically and creatively about model building

2. extrapolate from in-class exercises to simulate data and do inference with generalized linear models

3. hone your R skillz: function building, data visualization, etc.

\

### Question 1

Explain in your own words each of the terms in each of the equations below, 
as well as what each equation is doing, and what they are all doing together.

$Y_i \sim \mathcal{Pois}(\lambda_i=f(x_i))$ \
$\lambda = \text{exp}(f(x_i))$ \
$f(x_i) = \alpha + \beta x_i$ \

 - Y_i is the $i^{\text{th}}$ response variable
 - x_i is the $i^{\text{th}}$ predictor variable
 - $\lambda_i$ is the rate parameter of the Poisson distribution from which the $i^{\text{th}}$ datum is being modeled as a draw
 - $f$ is the function that relates the predictor variable to the expected value of its associated response variable (in this case, a linear model)

As a whole, these equations describe a Poisson regression with the canonical log link 
function and a linear deterministic function.

### Question 2

Write mle2 code blocks for the following models.
For each, define an R function for the inverse link function and 
the deterministic function, and write out the mle2 call 
using a data frame with named columns "predictor" and "response".

A) Binomial regression (sample size of 15)

```{r,eval=FALSE}
detFunc <- function(x,a,b){
	return(a + b*x)
}

invLink <- function(z){
	p <- 1/(1+exp(-z))
	return(p)
}

mod <- mle2(response ~ dbinom(size=15,
					   p=invLink(detFunc(predictor,a,b))
				 ),
			data=mydata,
			start=list("a"=0,"b"=0))
```

B) Poisson Regression

```{r,eval=FALSE}
detFunc <- function(x,a,b){
	return(a + b*x)
}

invLink <- function(z){
	return(exp(z))
}

mod <- mle2(response ~ dpois(lambda=invLink(
					   			detFunc(predictor,a,b)
					   	 	 )
				 	  ),
			data=mydata,
			start=list("a"=0,"b"=0))
```

C) Linear model (aka Normal regression)

```{r,eval=FALSE}
detFunc <- function(x,a,b){
	return(a + b*x)
}

invLink <- function(z){
	return(z)
}

mod <- mle2(response ~ dnorm(mean=invLink(
					   			detFunc(predictor,a,b),
					   			sd = s)
				 	  ),
			data=mydata,
			start=list("a"=0,"b"=0,"s"=1))
```


### Question 3

Use mle2 to infer the parameters of the model 
I used to simulate the data saved in the "mysteryData.Robj" 
file in the Class 24 homework folder on D2L.

Plot the line from the model you fit and also 
a histogram of the residuals (the observed data - 
their expected value for all data points). 
 
```{r,warning=FALSE}
detFunc <- function(a,b,x){
	return(a*x^b)
}

invLink <- function(z){
	return(z)
}

set.seed(124)
a <- 1.8
b <- 0.2
sigma <- 0.9
predictor <- runif(500,min=0,max=1000)
response <- rnorm(n=500,mean=invLink(detFunc(a,b,predictor)),sigma)

mysteryData <- data.frame("predictor"=predictor,"response"=response)
save(mysteryData,file="mysteryData.Robj")

library(bbmle)
mod <- mle2(response ~ dnorm(mean=invLink(detFunc(a,b,predictor)),sd=s),
			data=mysteryData,
			start=list("a"=rnorm(1),"b"=rnorm(1),"s"=abs(rnorm(1))))

tmp.x <- seq(0,1e3,length.out=1e3)
par(mfrow=c(1,2))
plot(mysteryData)
	lines(tmp.x,mod@coef[1]*tmp.x^mod@coef[2],col="red",lwd=2)
	
residuals <- mysteryData$response - mod@coef[1]* mysteryData$predictor^mod@coef[2]
hist(residuals)
```





