  ---
title: "GLM Class Activity"
output:
  html_document: default
language: R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
 First, install and load the bbmle library
```{r}
install.packages('bbmle')
library(bbmle)
```
 
Pitkow et al (1960) exposed guppies to 2, 8, 12, or 18 minutes in 5 degree Celsius water. The data – with a value of 1 for Alive meaning alive and a value of zero meaning dead. Use the code below to load their data.
 
```{r}
mydata <- read.csv('fishData.csv')
str(mydata)
 
```
 

1)
Below is code for running a linear normal regression GLM using mle2(). Try running the code and discuss with your group whether you think this model is doing a good job of fitting the data.
 
```{r}

#define a deterministic funciton
detFunc <- function(x,a,b){
    return(a + b*x)
}

# specify the inverse link function
invLink <- function(z){
    return(z)
}

# run a regression with mle2()
mod <- mle2(Alive ~ dnorm(mean=invLink( detFunc(Time,a,b)), sd=s),
            data=mydata,
            start=list("a"=0,"b"=0, 's'=1))

## use the model to generate data
modData = mod@coef[1]+mydata$Time*mod@coef[2]

## look at the model-predicted data curve on top of the actual data
plot(jitter(mydata$Time), jitter(mydata$Alive), bty="n", col="darkgray")
abline(mod, col="red")

residuals <- mydata$Alive - modData

plot(jitter(mydata$Time), jitter(residuals), bty="n", col="darkgray")
abline(h=0, lty=2)

```


2) 

Now, try a Binomial regression. Head to wikipedia and look up the link function for a Binomial regression: https://en.wikipedia.org/wiki/Generalized_linear_model#Link_function

Code for running a Binomial regression is below. I've added comments where things have changed from the previous version. Please discuss what has changed with your group.

```{r}

#define a deterministic function
detFunc <- function(x,a,b){
    return(a + b*x)
}

# specify the inverse link function
invLink <- function(z){
    return(exp(z)/(1+exp(z)))  ### CHANGE 
}

# run a regression with mle2()
mod2 <- mle2(Alive ~ dpois(lambda=invLink( detFunc(Time,a,b))), ### CHANGE 
            data=mydata,
            start=list("a"=0.1,"b"=0.1))  ##  CHANGE 


## use the model to generate predicted data
logOdds <- mod2@coef[1]+mydata$Time*mod2@coef[2]
mod2Data <- exp(logOdds)/(1+exp(logOdds)) ##CHANGE


## look at the model-predicted data curve on top of the actual data
plot(jitter(mydata$Time), jitter(mydata$Alive), bty="n", col="darkgray")
points(mydata$Time, mod2Data, col="red")


residuals2 <- mydata$Alive - mod2Data

plot(jitter(mydata$Time), jitter(residuals2), bty="n", col="darkgray")
abline(h=0, lty=2)

```

Discuss this Binomial regression model with your group. How is it different from the normal model? Which do you think fits better?
\
\
