---
title: "Final Project"
author: "Ankit Gupta"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario

We investigated how soil moisture levels influence plant growth, measured as plant height (cm). Soil moisture, ranging from 0% to 100%, was hypothesized to have a linear effect on height, with a baseline height of 2 cm at 0% moisture and an increase of 0.5 cm for every 1% rise in moisture. Observed heights were simulated using a linear model with added random noise (standard deviation = 1.5 cm) to reflect environmental variability, generating 1,000 data points for analysis.

```{r Simulation}

# Load required library
library(bbmle)

# Set seed for reproducibility
set.seed(123)

# Simulation Parameters
n <- 1000  # Number of data points
a <- 2     # Intercept
b <- 3     # Slope
sigma <- 1.5  # Standard deviation

# Simulate predictor values
predictor <- runif(n, 0, 10)  # Random values between 0 and 10

# Simulate response values
mu <- a + b * predictor  # Deterministic function (linear)
response <- rnorm(n, mean = mu, sd = sigma)  # Add random noise (Normal distribution)

# Combine into a data frame
simulated_data <- data.frame(predictor, response)

# Save to CSV for sharing
write.csv(simulated_data, "simulated_data_normal_linear.csv", row.names = FALSE)

# Plot the simulated data
plot(predictor, response, main = "Simulated Data (Normal with Linear Relationship)",
     xlab = "Predictor", ylab = "Response", pch = 16, col = "blue")


```


```{r MLE, echo=TRUE}

# Load required library
library(bbmle)

# Set seed for reproducibility
set.seed(123)

# Simulation Parameters
n <- 1000  # Number of data points
a <- 2     # Intercept (baseline height when soil moisture = 0)
b <- 0.5   # Slope (increase in height per unit of soil moisture)
sigma <- 1.5  # Standard deviation (random noise)

# Simulate soil moisture values (predictor)
soil_moisture <- runif(n, 0, 100)  # Soil moisture percentage (0 to 100)

# Simulate plant heights (response)
expected_height <- a + b * soil_moisture  # Deterministic function
plant_height <- rnorm(n, mean = expected_height, sd = sigma)  # Add normal noise

# Combine into a data frame
simulated_data <- data.frame(soil_moisture, plant_height)

# Plot the simulated data
plot(soil_moisture, plant_height, main = "Simulated Data (Linear Relationship)",
     xlab = "Soil Moisture (%)", ylab = "Plant Height (cm)", pch = 16, col = "blue")

# Define the negative log-likelihood function
nll <- function(a, b, sigma) {
  expected_height <- a + b * simulated_data$soil_moisture  # Deterministic function
  -sum(dnorm(simulated_data$plant_height, mean = expected_height, sd = sigma, log = TRUE))  # Log-likelihood
}

# Fit the model using MLE
mle_model <- mle2(
  minuslogl = nll,
  start = list(a = 1, b = 0.1, sigma = 1),  # Initial parameter guesses
  method = "L-BFGS-B",  # Optimization method
  lower = c(a = -Inf, b = -Inf, sigma = 0.01)  # Constraints (e.g., sigma > 0)
)

# Print the summary of the MLE fit
summary(mle_model)

# Extract estimated parameters
estimated_params <- coef(mle_model)
cat("Estimated Parameters:\n")
print(estimated_params)

# Add the fitted line to the plot
lines(simulated_data$soil_moisture,
      estimated_params["a"] + estimated_params["b"] * simulated_data$soil_moisture,
      col = "red", lwd = 2)

# Add a legend to the plot
legend("topleft", legend = c("Simulated Data", "Fitted Line"),
       col = c("blue", "red"), pch = c(16, NA), lwd = c(NA, 2))


```


```{r csv}

# Save the simulated data to a CSV file
write.csv(simulated_data, "simulated_seed_germination_data.csv", row.names = FALSE)

# Print a message to confirm the file is saved
cat("Simulated data has been saved to 'simulated_seed_germination_data.csv' in the working directory.\n")


```